\begin{thebibliography}{99}

\bibitem[Bma06]{b:bmad}
D. Sagan,
"Bmad: A Relativistic Charged Particle Simulation Library"
Nuc.\ Instrum.\ \& Methods Phys.\ Res.\ A, {\bf 558}, pp 356-59 (2006).

The Bmad Manual can be optained at:
\hfill\break
\hspace*{20pt} 
\url{www.classe.cornell.edu/bmad}

\bibitem[Bengt97]{b:bengtsson}
J.~Bengtsson,
``The Sextupole Scheme for the Swiss Light Source (SLS): An Analytic Approach,''
SLS Note 9/97, Paul Scherrer Institut, (1997).

\bibitem[Blender]{b:blender}
\vn{Blender} web page:
\hfill\break
\hspace*{20pt} 
\url{blender.org/}

\bibitem[Fra11]{b:emit}
A.~Franchi, L.~Farvacque, J.~Chavanne, F.~Ewald, B.~Nash, K.~Scheidt, and R.~Tom\',
``Vertical emittance reduction and preservation in electron storage rings via resonance driving terms correction'',
Phys. Rev. ST Accel. Beams,
{\bf 14}, 3, 034002, (2011). 
\hfill\break
\hspace*{20pt}
\url{link.aps.org/doi/10.1103/PhysRevSTAB.14.034002}

\bibitem[CBETA19]{b:cbeta}
C.~Gulliford, A.~Bartnik, J.~Scott~Berg, J.~Dobbins, A.~Nunez-delPrado, and D.~Sagan,
``Experience With CBETA Online Modeling Tools'',
13th International Computational Accelerator Physics Conference, 2019.

\bibitem[NR92]{b:nr}
W.~Press, B.~Flannery, S.~Teukolsky, W.~Wetterling,
{\em Numerical Recipes in Fortran, the Art of Scientific Computing},
Second Edition, Cambridge University Press, New York (1992)

\bibitem[Montague]{b:w}
B. W. S. L. Montague, “Linear Optics For Improved Chromaticity Correction,” 
CERN, Geneva, Tech. Rep. CERN-LEP-Note-165. LEP-Note-165, 1979. 
\url{http://cds.cern.ch/record/443342}

\bibitem[Saf97]{b:orm}
J. Safranek, ``Experimental determination of storage ring optics
using orbit response measurements'', NIM-A388, p. 27 (1997).

\bibitem[Sag99]{b:linear.coupled}
D. Sagan, and D. Rubin,
``Linear Analysis of Coupled Lattices,''
Phys. Rev. ST Accel. Beams 2, 074001 (1999).
\hfill\break
\hspace*{20pt}
\url{https://journals.aps.org/prab/abstract/10.1103/PhysRevSTAB.2.074001}

\bibitem[Sag00a]{b:beta.meas}
D. Sagan, R. Meller, R. Littauer, and D. Rubin,
``Betatron phase and coupling measurement at the Cornell Electron/Positron Storage Ring'', 
Phys. Rev. ST Accel. Beams 3, 092801 (2000).
\hfill\break
\hspace*{20pt}
\url{link.aps.org/doi/10.1103/PhysRevSTAB.3.092801}

\bibitem[Sag00b]{b:wave}
D. Sagan,
``Betatron phase and coupling correction at the Cornell Electron/Positron
Storage Ring'', Phys. Rev. ST Accel. Beams 3, 102801 (2000).
\hfill\break
\hspace*{20pt}
\url{link.aps.org/doi/10.1103/PhysRevSTAB.3.102801}

\bibitem[Sto96]{b:de}
R.~Storn, and K.~V.~Price, "Minimizing the real function of the
ICEC'96 contest by differential evolution" IEEE conf. on Evolutionary
Computation, 842-844 (1996).

\bibitem[Wang12]{b:wang}
C.~Wang,
``Explicit formulas for 2nd-order driving terms due to sextupoles and
chromatic effects of quadrupoles,''
ANL/APS/LS-330, Argonne National Laboratory, (2012).

\bibitem[Wil00]{b:wille}
Klaus Wille, {\em The Physics of Particle Accelerators: An Introduction},
Translated by Jason McFall, Oxford University Press (2000).

\end{thebibliography}
