\chapter{Organization and Structure}
\label{c:organization}

This chapter discusses how \tao is organized. After you are familiar with the basics of \tao, you
might be interested to exploit its versatility by extending \tao to do custom calculations. For
this, see Chapter~\ref{c:custom.tao}.

%----------------------------------------------------------------
\section{The Organization of Tao: The Super\_Universe}
\label{s:organization}
\index{super_universe}

Many simulation problems fall into one of three categories: 
\begin{itemize}
\item 
Design a lattice subject to various constraints.
\item
Simulate errors and changes in machine parameters. For example, you want to simulate what happens to
the orbit, beta function, etc., when you change something in the machine.
\item 
Simulate machine commissioning including simulating data measurement and correction. For example,
you want to know what steering strength changes will make an orbit flat.
\end{itemize}
Programs that are written to solve these types of problems have common elements: You have variables
you want to vary in your model of your machine, you have "data" that you want to view, and, in the
first two categories above, you want to match the machine model to the data (in designing a lattice
the constraints correspond to the data).

With this in mind, \tao was structured to implement the essential ingredients needed to solve these
simulation problems.  The information that \tao knows about can be divided into five (overlapping)
categories:
\begin{description}
  \index{lattice}
  \item[Lattice] \Newline   
Machine layout and component strengths, and the beam orbit (\sref{s:lattice}).
  \index{data}
  \item[Data] \Newline
Anything that can be measured.
For example: The orbit of a particle or the lattice beta 
functions, etc. (\sref{c:data})
  \index{variable}
  \item[Variables] \Newline
Essentially, any lattice parameter or initial condition that can be varied. For example: quadrupole
strengths, etc. (\sref{c:var}).
  \index{plotting}
  \item[Plotting]  \Newline
Information used to draw graphs, display the lattice 
floor plan, etc. (\sref{c:plotting}).
  \index{global parameters}
  \item[Global Parameters] \Newline
 \tao has a set of parameters to control every aspect of how it behaves from
the random number seed \tao uses to what optimizer is used for fitting data.
\end{description}

%------------------------------------------------------------------------
\section{The Super\_universe}
\label{s:super.uni}
\index{super_universe|hyperbf}

\index{structure|hyperbf}
The information in \tao deals is organized in a hierarchy of \vn{``structures''}. At the top level,
everything known to \tao is placed in a single structure called the \vn{super_universe}.

\index{universe}
\index{variable}
Within the \vn{super_universe}, lies one or more \vn{universes} (\sref{s:universe}), each
\vn{universe} containing a particular machine lattice and its associated data. This allows for the
user to do analysis on multiple machines or multiple configurations of a single machine at the same
time. The \vn{super_universe} also contains the \vn{variable}, \vn{plotting}, and \vn{global
parameter} information.

%------------------------------------------------------------------------
\section{The Universe}
\label{s:universe}
\index{universe|hyperbf}

\index{lattice}\index{design lattice}\index{model lattice}
\index{base lattice}\index{data}\index{super_universe}
The \tao \vn{super_universe} (\sref{s:super.uni}) contains one or more \vn{universes}.  A
\vn{universe} contains a \vn{lattice} (\sref{s:lattice}) plus whatever data (\sref{c:data}) one
wishes to study within this lattice (i.e. twiss parameters, orbit, phase, etc.). Actually, there are
three lattices within each universe: the \textbf{design} lattice, \textbf{model} lattice and
\textbf{base} lattice. Initially, when \tao is started, all three lattices are identical and
correspond to the lattice read in from the lattice description file (\sref{s:init.lat}).

There are several situations in which multiple universes are useful. One case where multiple
universes are useful is where data has been taken under different machine conditions. For example,
suppose that a set of beam orbits have been measured in a storage ring with each orbit corresponding
to a different steering element being set to some non-zero value. To determine what quadrupole
settings will best reproduce the data, multiple universes can be setup, one universe for each of the
orbit measurements. Variables can be defined to simultaneously vary the corresponding quadrupoles in
each universe and \tao's built in optimizer can vary the variables until the data as determined from
the \vn{model} lattice (\sref{s:lattice}) matches the measured data. This \vn{orbit response matrix}
(ORM) analysis is, in fact, a widely used procedure at many laboratories.

If multiple universes are present, it is important to be able to specify, when issuing commands to
tao and when constructing \tao initialization files, what universe is being referred to when
referencing parameters such as data, lattice elements or other stuff that is universe
specific. [Note: \tao variables are {\em not} universe specific.] If no universe is specified with a
command, the \vn{default} universe will be used. This default universe is set set by the \vn{set
default universe} command (\sref{s:set}). When \tao starts up, the default universe is initially set
to universe 1. Use the \vn{show global} (\sref{s:show}) command to see the current default universe.

the syntax used to specify a particular universe or range of universes is attach a prefix of the
form:
\begin{example}
  [<universe_range>]@<parameter>
\end{example}
Commas and colons can be used in the syntax for \vn{<universe_range>}, similar to the \vn{element
list} format used to specify lattice elements (\sref{s:ele.list.format}).  When there is only a
single Universe specified, the brackets \vn{[...]} are optional. When the universe prefix is not
present, the current default universe is used. The current default universe can also be specified
using the number \vn{-1}. Additionally, a ``\vn{*}'' can be used as a wild card to denote all of the
universes. Examples:
\begin{example}
  [2:4,7]@orbit.x ! The \vn{orbit.x} data in universes 2, 3, 4 and 7.
  [2]@orbit.x     ! The \vn{orbit.x} data in universe 2. 
  2@orbit.x       ! Same as "2@orbit.x".
  orbit.x         ! The \vn{orbit.x} data in the current default universe.
  -1@orbit.x      ! Same as "orbit.x".
  *@orbit.x       ! orbit.x data in all the universes.
  *@*             ! All the data in all the universes. 
\end{example}

%------------------------------------------------------------------------
\section{Lattices}
\index{lattice|hyperbf}
\label{s:lattice}

\index{design lattice}\index{model lattice}
\index{base lattice}
A \vn{lattice} consists of a machine description (the strength and placement of elements such as
quadrupoles and bends, etc.), along with the beam orbit through them. There are actually three types
of lattices:
  \vspace*{-3ex}
  \begin{description}
  \index{design lattice|hyperbf}
  \item[Design Lattice] \Newline 
The \vn{design} lattice corresponds to the lattice read in from the lattice description file(s)
(\sref{s:init.lat}). In many instances, this is the particular lattice that one wants the actual
physical machine to conform to. The \vn{design} lattice is fixed. Nothing is allowed to vary in this
lattice.
  \index{model lattice|hyperbf}
  \item[Model Lattice] \Newline
Initially the \vn{model} lattice is the same as the \vn{design} lattice. Except for some commands
that explicitly set the \vn{base} lattice, all \tao commands to vary lattice variables vary
quantities in the \vn{model} lattice. In particular, things like orbit correction involve varying
\vn{model} lattice variables until the \vn{data}, as calculated from the \vn{model}, matches the
\vn{data} as actually measured.
  \index{base lattice|hyperbf}
  \index{base lattice!using set command}
  \item[Base Lattice] \Newline
It is sometimes convenient to designate a reference lattice so that changes in the \vn{model} from
the reference point can be examined.  This reference lattice is called the \vn{base} lattice. The
\vn{set} command (\sref{s:set}) is used to transfer information from the \vn{design} or \vn{model}
lattices to the base lattice. Initially, the \vn{base} lattice is set equal to the \vn{design}
lattice by \tao.
  \end{description}

Lattices can have multiple \vn{branches}. For example, two intersecting rings can be represented as
a lattice with two branches, one for each ring. See the \bmad manual for more details. Many \tao
commands operate on a particular lattice branch. For example, the \vn{show lat} command prints the
lattice elements of a particular branch. If no branch is specified with a command, the default
branch is used. The default branch is set with the \vn{set default branch} command
(\sref{s:set}). Initially, when \tao is started, the default branch is set to branch 0. Use the
\vn{show global} (\sref{s:show}) command to see the current default branch.

%------------------------------------------------------------------------
\section{Tracking Types}
\index{tracking!types}

\index{track_type}
\index{tao_global_struct}
\index{global%track_type}
The are two types of tracking implemented in \tao: single particle tracking and many particle
multi-bunch tracking.  Single particle tracking is just that, the tracking of a single particle
through the lattice. Many particle multi-bunch tracking creates a Gaussian distribution of particles
at the beginning of the lattice and tracks each particle through the lattice, including any
wakefields.  Single particle tracking is used by default. The \vn{global%track_type} parameter
(\sref{s:globals}), which is set in the initialization file, is used to set the tracking.

Particle spin tracking has also been set up for single particle and many particle tracking. See
Sections~\sref{s:globals} and \sref{s:beam.init} for details on setting up spin tracking.

%------------------------------------------------------------------------
\section{Lattice Calculation}\index{lattice!calculation of}
\label{s:lat.calc}

After each \tao command is processed, the lattice and ``merit'' function are recalculated and the
plot window is regenerated. The merit function determines how well the \vn{model} fits the measured
data. See Chapter~\ref{c:opti} for more information on the merit function and its use by the
optimizer.

Below are the steps taken after each \tao command execution:
\begin{enumerate}
  \item 
The data and variables used by the optimizer are re-determined. This is affected by commands such as
\vn{use, veto,} and \vn{restore} and any changes in the status of elements in the ring (e.g. if any
elements have been turned off).
  \item 
If changes have been made to the lattice (e.g. variables changed) then the model lattice for all
universes will be recalculated. The \vn{model} orbit, linear transfer matrices and Twiss parameters
are recalculated for every element. All data types will also be calculated at each element specified
in the initialization file.  For single particle tracking the linear transfer matrices and Twiss
parameters are found about the tracked orbit. Tracking is performed using the tracking method
defined for each element (i.e. Bmad Standard, Symplectic Lie, etc...). See the \bmad Reference
manual for details on tracking and finding the linear transfer matrices and Twiss parameters.
  \item 
The \vn{model} data is recalculated from the \vn{model} orbit, linear transfer matrices, Twiss
parameters, particle beam information and global lattice parameters.  Any custom data type
calculations are performed \textit{before} the standard \tao data types are calculated.
  \item 
Any user specified data post-processing is performed in
\vn{tao_hook_post_process_data}.
  \item 
The contributions to the merit function from the variables and data are computed.
  \item 
Data and variable values are transferred to the plotting structures.
  \item 
The plotting window is regenerated.
\end{enumerate}

If a closed orbit is to be calculated, \tao uses an iterative method to converge on a solution where
\tao starts with some initial orbit at the beginning of the lattice, tracks from this initial orbit
through to the end of the lattice, and then adjusts the beginning orbit until the end orbit matches
the beginning orbit. A problem arises if the tracked particle is lost before it reaches the end of
the lattice since \tao has no good way to calculate how to adjust the beginning orbit to prevent the
particle from getting lost. In this case, \tao, in desperation, will try the orbit specified by
\vn{particle_start} in the \bmad lattice file (see the \bmad manual for more details on setting
\vn{particle_start}). Note: \vn{particle_start} can be varied while running \tao using the \vn{set
particle_start} (\sref{s:set}) or \vn{change particle_start} (\sref{s:change}) commands.

If the recalculation takes a significant amount of time, the recalculation may be suppressed using
the \vn{set global lattice_calc_on} command (\sref{s:set.global}) or the \vn{set universe} command
(\sref{s:set.universe}).
