\chapter{Python Interface to Tao}
\index{python interface}
\label{c:python}

It is sometimes convenient to interface \tao to a scripting language like Python or interface to
some external program.  Applications include analyzing Tao generated data or 
to interface \tao to an online control system environment.

To aid in interfacing, \tao has the \vn{pipe} command (\sref{s:pipe}).\footnote
  {
Formally this command was called \vn{python} but the name was changed to avoid confusion with
the scripting language Python.
  }
The \vn{pipe} command defaines a standardized syntax with which to communicate with \tao.

Another aid is the \vn{PyTao} package which is an interface layer to be used between \tao and
\vn{Python}. See \sref{s:pytao} for more details.

%--------------------------------------------------------------------------
\section{PyTao Interface}
\label{s:pytao}

The \vn{PyTao} package is an interface layer to be used between \tao and
\vn{Python}. \vn{PyTao} is hosted on \vn{GitHub}
(independent of \bmad distributions) at:
\vspace{-2ex}
\begin{itemize}
  \item[] \url{https://bmad-sim.github.io/pytao}
\end{itemize}
\vspace{-2ex}
Documentation for setup and using PyTao is at:
\begin{example}
  bmad-sim.github.io/pytao/
\end{example}
See the \vn{PyTao} documentation for installation instructions, examples, etc. In this chapter, some
simple examples will be given.

The \vn{PyTao} package uses \tao's \vn{pipe} command to ease integration with \vn{Python}.

There are two ways to interface with Python/PyTao. One way is using the Python \vn{ctypes}
library. The other way is using the \vn{pexpect} module. A Web search will point to documentation on
\vn{ctypes} and \vn{pexpect}. 

\vn{ctypes} is a foreign function library for Python which can be used to link to a \tao shared
library.  The \vn{pexpect} module is a general purpose tool for interfacing Python with programs
like \tao. If \vn{pexpect} is not present your system, it can be downloaded from
\vn{www.noah.org/wiki/pexpect}.

The advantage of \vn{ctypes} is that it directly accesses \tao code
which makes communication between Python and \tao more robust. The disadvantage of \vn{ctypes} is
that it needs a shared-object version of the \vn{Tao} library. [See the Bmad web site for
information on building shared-object libraries.] The disadvantage of \vn{pexpect} is that it is
slower and it is possible for \vn{pexpect} to time out waiting for a response from \tao.

%--------------------------------------------------------------------------
\subsection{Python/PyTao Via Pexpect}

For communicaiton via \vn{pexpect} (\sref{s:pytao}), the python module \vn{tao_pipe.py}, is
provided by \vn{PyTao} in the directory \vn{pytao/tao_pexpect}.

Example Python session:
\begin{example}
  >>> from pytao.tao_pexpect import tao_pipe  # import module
  >>> p = tao_pipe.tao_io("-lat my_lat.bmad") # init session
  >>> out = p.cmd_in("show global")           # Command to Tao
  >>> print(out)                              # print the output from Tao
  >>> p.cmd("show global")                    # Like p.cmd_in() excepts prints the output too.
\end{example}

%--------------------------------------------------------------------------
\subsection{Python/PyTao Interface Via Ctypes}

A \vn{ctypes} based Python module \vn{pytao.py} for interfacing \tao to \vn{Python} is provided by
\vn{PyTao} (\sref{s:pytao}) in the directory \vn{pytao/tao_pexpect}.

A test driver script named \vn{pytao_example.py} is in the same directory. See the documentation in
both of these files for further information.

%--------------------------------------------------------------------------
\section{Tao's Pipe Command}
\label{s:pipe}

\tao's \vn{pipe} (\sref{s:pipe}) command was developed to:
%
\begin{itemize}
\item 
Standardize output of information (data, parameters, etc.) from \tao to simplify the task of
interfacing \tao to external programs especially scripting languages like \vn{Python}.
% 
\item 
Act as an intermediate layer for the control of \tao by such things as machine online control
programs or the planned graphical user interface for \tao.
\end{itemize}

Using the \vn{pipe} command to control \tao will not be covered here. The interested reader is
invited to read the sections of this manual on the coding of \tao and look at the \tao code itself
(which is heavily documented).

Using the \vn{pipe} command is far superior to using the \vn{show} command when interfacing to an
external program. For one, the \vn{pipe} command is formatted for ease of parsing. Another reason
to use the \vn{pipe} command is that, as \tao is developed over time, the output format of the
\vn{pipe} command is much more stable than output from the \vn{show} command.\footnote
  {
The output of the \vn{pipe} command will change when \tao's or \bmad's internal structures are
modified. This is in contrast to the \vn{show} command whose output is formated to be human readible
and whose output format may change on a whim.
  }
Thus the risk of User developed interface code breaking is much reduced by using the \vn{pipe} command.

The general form of the \vn{pipe} command is:
\begin{example}
  pipe <subcommand> <arguments>
\end{example}
The \vn{pipe} command has a number of \vn{subcommands} (over 100) that are listed in
\Sref{s:pipe.sub}. The sub-commands can be divided into two categories. One category are the
``\vn{action}'' subcommands which allow the user to control \tao (for example, creating variables
and data for use in an optimization). The other category are the ``\vn{output}'' subcommands which
output information from \tao.

The output of the \vn{pipe} command are semi-colon delimited lists. Example: With the
\vn{pipe global} the output looks like:
\begin{example}
  lm_opt_deriv_reinit;REAL;T; -1.0000000000000000E+00
  de_lm_step_ratio;REAL;T;  1.0000000000000000E+00
  de_var_to_population_factor;REAL;T;  5.0000000000000000E+00
  unstable_penalty;REAL;T;  1.0000000474974513E-03
  n_opti_cycles;INT;T;20
  track_type;ENUM;T;single
  derivative_uses_design;LOGIC;T;F
  ... etc ...
\end{example}

Most \vn{output} subcommands use ``\vn{parameter list form}'' format where each line has four fields
separated by semicolons:
\begin{example}
  {name};{type};{variable};{value(s)}
\end{example}
The fields are:
\begin{example}
    name:       The name of the parameter

    type:       The type of the parameter:
        INT           Integer number
        REAL          Real number
        COMPLEX       Complex number. A complex number is output as Re;Im
        REAL_ARR      Real array
        LOGIC         Logical: "T" or "F".
        INUM          Integer whose allowed values can be obtained 
                        using the "pipe inum" command.
        ENUM          String whose allowed values can be obtained 
                        using the "pipe enum" command.
        FILE          Name of file.
        CRYSTAL       Crystal name string. EG: "Si(111)"
        DAT_TYPE      Data type string. EG: "orbit.x"
        DAT_TYPE_Z    Data type string if plot%x_axis_type = 'data'. 
                        Otherwise is a data_type_z enum.
        SPECIES       Species name string. EG: "H2SO4++"
        ELE_PARAM     Lattice element parameter string. EG "K1"
        STR           String that does not fall into one of the above string categories.
        STRUCT        Structure. In this case {component_value(s)} is of the form:
                        {name1};{type1};{value1};{name2};{type2};{value2};...
        COMPONENT     For curve component parameters.

    can_vary:   Either 'T', 'F', or 'I', indicating whether or not the
                user may change the value of the parameter. 'I' indicates
                that the parameter is to be ignored by a GUI when displaying parameters.

    value(s):   The value or values of the the parameter. If a parameter has multiple
                values (EG an array), the values will be separated by semicolons.
\end{example}

%--------------------------------------------------------------------------
\section{Plotting Issues}
\label{s:gui.plot}

When using \tao with a \vn{GUI}, and when the \vn{GUI} is doing the plotting, the \vn{-noplot} and
\vn{-external_plotting} options (\sref{s:command.line}) should be used when starting \tao. The
\vn{-noplot} option (which sets \vn{global%plot_on}) prevents \tao from opening a plotting
window. Note: Both of these options can also be set, after startup, with the \vn{set global} command
and the setting of both can be viewed using the \vn{show global} command.

With \vn{-external_plotting} set, the external code should handle how plots are assigned to plot
regions and it would be potentially disruptive if a user tired to place plots (which could
inadvertently happen when running command files). To avoid this, with \vn{-external_plotting} set,
the \vn{place} command will not do any placement but rather save the \vn{place} arguments (which is
the name of a template plot and a region name) to a buffer which then can be read out by the
external code using the \vn{pipe place_buffer} command. The external code may then decide how to
proceed. The external code is able to bypass the buffering and perform placements by using
\vn{place} with the \vn{-no_buffer} switch (\sref{s:place}). Notice: \tao never processes place
command information put in the buffer. It is up to the external code to decide on a course of action.

Normally when \tao is not displaying the plot page when the \vn{-noplot} option is used, \tao will,
to save time, not calculate the points needed for plotting curves. The exception is if
\vn{-external_plotting} is turned on. In this case, to make plot references unambiguous, plot can be
referred to by their index number. The plot index number can be viewed using the \vn{pipe
plot_list} command. Template plots can be referenced using the syntax ``\vn{@Tnnn}'' where \vn{nnn}
is the index number. For example, \vn{@T3} referrers to the template plot with index 3. Similarly,
the displayed plots (plots that are associated with plot regions) can be referred to using the
syntax ``\vn{@Rnnn}''.

%--------------------------------------------------------------------------
\section{Pipe subcommands}
\label{s:pipe.sub}

The \vn{pipe} command has the following subcommands:

\input{pipe-interface-commands.tex}

