\chapter{Beam Initialization}
\label{c:beam.init}

Some \bmad based programs track beams of particles instead of tracking individual particles
one-by-one. This can be useful for several reasons. For example, tracking beams is useful when
inter-bunch or intra-bunch effects are to be simulated. Also tracking beams can simplify the
bookkeeping a program needs to do to calculate such quantities such as the bunch size.

A \bmad based program has two standard ways to specify the initial distribution of a beam. One is
using a \vn{beam_init_struct} \vn{structure} (\sref{s:struct}) which holds parameters (for
example, the beam emittances) from which a distribution of particles can be constructed. The
\vn{beam_init_struct} structure is explained in Section~\sref{s:beam.init.struct}. The other way is to
specify the initial beam distribution via a file that has the individual particle positions. This is
covered in Section~\sref{s:beam.init.file}.

%-----------------------------------------------------------------
\section{Beam_Init_Struct Structure}
\label{s:beam.init.struct}
\index{beam initialization parameters|hyperbf}

\index{beam_init_struct|hyperbf}
The \vn{beam_init_struct} \vn{structure} (\sref{s:struct}) holds parameters which are used to
initialize a beam. The parameters of this structure, shown with their default values, are:
\begin{example}
  type beam_init_struct
    character(200) :: position_file = ""       ! Initialization file name.
    character distribution_type(3)             ! "ELLIPSE", "KV", "GRID", "" (default).
    type (ellipse_beam_init_struct) ellipse(3) ! For ellipse beam distribution
    type (kv_beam_init_struct) KV              ! For KV beam distribution
    type (grid_beam_init_struct) grid(3)       ! For grid beam distribution
    logical use_particle_start = F     ! Use particle_start instead of %center and %spin?
    character random_engine            ! "pseudo" (default) or "quasi". 
    character random_gauss_converter   ! "exact" (default) or "quick". 
    real center(6) = 0                 ! Bench phase space center offset.
    real t_offset = 0                  ! Time offset.
    real center_jitter(6) = 0          ! Bunch center rms jitter
    real emit_jitter(2)   = 0          ! %RMS a and b-mode emittance jitter
    real sig_z_jitter     = 0          ! bunch length RMS jitter 
    real sig_pz_jitter     = 0         ! pz energy spread RMS jitter 
    real random_sigma_cutoff = -1      ! -1 => no cutoff used.
    integer n_particle = 0             ! Number of simulated particles per bunch.
    logical renorm_center = T          ! Renormalize centroid?
    logical renorm_sigma = T           ! Renormalize sigma?
    real spin(3) = 0, 0, 0             ! Spin (x, y, z)
    real a_norm_emit = 0               ! a-mode normalized emittance (= \(\beta\,\gamma\,\epsilon_{unnorm}\))
    real b_norm_emit = 0               ! b-mode normalized emittance (= \(\beta\,\gamma\,\epsilon_{unnorm}\))
    real a_emit = 0                    ! a-mode unnormalized emittance (= \(\epsilon_{norm}/\beta\gamma\))
    real b_emit = 0                    ! b-mode unnormalized emittance (= \(\epsilon_{norm}/\beta\gamma\))
    real dpz_dz = 0                    ! Correlation of pz with longitudinal position.
    real dt_bunch = 0                  ! Time between bunches.
    real sig_z = 0                     ! Z sigma in m.
    real sig_pz = 0                    ! pz sigma.
    real bunch_charge = 0              ! Charge in a bunch.
    integer n_bunch = 0                ! Number of bunches.
    integer ix_turn = 0                ! Turn index.
    character species = ""             ! Species. Default is reference particle.
    logical full_6D_coupling_calc = F  ! Use 6x6 1-turn matrix to match distribution?  
    logical use_t_coords = F  ! If true, the distributions will be 
                              !   calculated using time coordinates (\sref{s:time.phase.space}). 
    logical use_z_as_t   = F  ! Only used if use_t_coords = T:
                              !   If True,  the z coordinate stores the time.
                              !   If False, the z coordinate stores the s-position.
  end type
\end{example}
Note: The \vn{z} coordinate value given to particles of a bunch is with respect to the
nominal center of the bunch. Therefore, if there are multiple bunches, and there is an RF
cavity whose frequency is not commensurate with the spacing between bunches, absolute time
tracking (\sref{s:rf.time}) must be used.

\begin{description}
\item[\%position_file] \Newline
\vn{%position_file} sets the name of the file to be read in containing the particle coordinates.  Input
from a file is triggered if not-blank. The format of the file is discussed in
Section~\sref{s:beam.init.file}.
%
\item[\%a_emit, \%b_emit, \%a_norm_emit, \%b_norm_emit] \Newline
Normalized and unnormalized emittances. Either \vn{a_norm_emit} or \vn{a_emit} may be set but not
both. similarly, either \vn{b_norm_emit} or \vn{b_emit} may be set but not both.

When simulating a ring, if any of these parameters is set negative, and if the \bmad based program
being run has enabled it, the value of the parameter will be set the value as calculated from the
lattice using synchrotron radiation integral formulas (\sref{s:synch.ints}).
%
\item[\%bunch_charge] \Newline
The \vn{%bunch_charge} paramter sets the charge of a bunch. If reading from a file, the bunch
charge will be set to the value of \vn{%bunch_charge} except if \vn{%bunch_charge} has a value
of zero in which case the bunch charge as specified in the file is used.
%
\item[\%center(6)] \Newline
The \vn{%center} parameter is used to offset the center position of a bunch. Exception: When
\vn{%use_particle_start} is set to True, the \vn{particle_start} orbital values are used instead for the
center position offset. See the description of \vn{%use_particle_start} below for more details.
%
\item[\%center_jitter, \%emit_jitter, \%sig_z_jitter, \%sig_pz_jitter] \Newline
These components can be used to provide a bunch-to-bunch random variation in the emittance and bunch
center.
%
\item[\%distribution_type(3)] \Newline
The \vn{%distribution_type(:)} array determines what algorithms are used to generate
the particle distribution for a bunch. Note: If \vn{%position_file} is not blank, the
beam distribution will be read from the appropriate file and \vn{%distribution_type}
will be ignored.

\vn{%distribution_type(1)} sets the distribution 
type for the $(x, p_x)$ 2D phase space, etc. 
Possibilities for \vn{%distribution_type(:)} are:
\begin{example}
  "", or "RAN_GAUSS"  ! Random distribution (default).
  "ELLIPSE"           ! Ellipse distribution (\sref{s:ellipse.init})
  "KV"                ! Kapchinsky-Vladimirsky distribution (\sref{s:kv.init})
  "GRID"              ! Uniform distribution.
\end{example}
Since the Kapchinsky-Vladimirsky distribution is for a 4D phase space, if the Kapchinsky-Vladimirsky
distribution is used, \vn{"KV"} must appear exactly twice in the \vn{%distribution_type(:)} array.

Unlike all other distribution types, the \vn{GRID} distribution is independent of the Twiss
parameters at the point of generation.  For the non-\vn{GRID} distributions, the distributions are
adjusted if there is local $x$-$y$ coupling (\sref{s:coupling}). For lattices with a closed
geometry, if \vn{full_6D_coupling_calc} is set to \vn{True}, the full 6-dimensional coupling matrix
is used. If \vn{False}, which is the default, The 4-dimensional $\bfV$ matrix of \Eq{vgicc1} is
used.

Note: The total number particles generated is the product of the individual
distributions. For example:
\begin{example}
  type (beam_init_struct) bi
  bi%distribution_type = ELLIPSE", "ELLIPSE", "GRID"
  bi%ellipse(1)%n_ellipse = 4
  bi%ellipse(1)%part_per_ellipse = 8
  bi%ellipse(2)%n_ellipse = 3
  bi%ellipse(2)%part_per_ellipse = 100
  bi%grid(3)%n_x = 20
  bi%grid(3)%n_px = 30
\end{example}
The total number of particles per bunch will be $32 \times 300 \times 600$. The exception is that
when \vn{RAN_GAUSS} is mixed with other distributions, the random distribution is overlaid with the
other distributions instead of multiplying. For example:
\begin{example}
  type (beam_init_struct) bi
  bi%distribution_type = RAN_GAUSS", "ELLIPSE", "GRID"
  bi%ellipse(2)%n_ellipse = 3
  bi%ellipse(2)%part_per_ellipse = 100
  bi%grid(3)%n_x = 20
  bi%grid(3)%n_px = 30
\end{example}
Here the number of particle is $300 \times 600$. Notice that when \vn{RAN_GAUSS} is mixed with other
distributions, the value of \vn{beam_init%n_particle} is ignored.
%
\item[\%dPz_dz] \Newline
Correlation between $p_z$ and $z$ phase space coordinates. 
%
\item[\%dt_bunch] \Newline
Time between bunches
%
\item[\%ellipse(3)] \Newline
The \vn{%ellipse(:)} array sets the parameters for the 
\vn{ellipse} distribution (\sref{s:ellipse.init}). 
Each component of this array looks like
\begin{example}
  type ellipse_beam_init_struct
    integer part_per_ellipse  ! number of particles per ellipse.
    integer n_ellipse         ! number of ellipses.
    real sigma_cutoff         ! sigma cutoff of the representation.
  end type
\end{example}
%
\item[\%full_6D_coupling_calc] \Newline
If set \vn{True}, coupling between the transverse and longitudinal modes is taken into
account when calculating the beam distribution. 
The default \vn{False} decouples the transverse and longitudinal calculations.
%
\item[\%grid(3)] \Newline
The \vn{%grid} component of the \vn{beam_init_struct} sets the parameters 
for a uniformly spaced grid of particles. The components of \vn{%grid}
are:
\begin{example}
  type grid_beam_init_struct
    integer n_x        ! number of columns.
    integer n_px       ! number of rows.
    real x_min         ! Lower x limit.
    real x_max         ! Upper x limit.
    real px_min        ! Lower px limit.
    real px_max        ! Upper px limit.
  end type
\end{example}
%
\item[\%ix_turn] \Newline
Turn index. This affects how particle time is calculated. Particle time is calculated from phase
space $z$ and $t_0$ via \Eq{zbctt}. $t_0$ is the reference time at the lattice element that the beam
is initialized at (\sref{s:ref.energy}). For simulations where the beam is circulating in a ring
over many turns, it may be desired to initialize the beam appropriate for some turn after the first
turn. That is, set the particle time with the reference time $t_{0n}$ associated with the $n$\Th
(set by \vn{ix_turn}) turn at which the beam is being initialized at
\begin{equation}
  t_{0n} = t_0 + n \, t_{rev} 
\end{equation}
where $t_{rev}$ is the revolution time.
%
\item[\%KV] \Newline
The \vn{%kv} component of the \vn{beam_init_struct} sets the parameters for the 
Kapchinsky-Vladimirsky distribution (\sref{s:kv.init}). The components of \vn{%KV}
are:
\begin{example}
  type kv_beam_init_struct
    integer part_per_phi(2)    ! number of particles per angle variable.
    integer n_I2               ! number of I2
    real A                     ! A = I1/e
  end type
\end{example}
%
\item[\%n_bunch] \Newline
The number of bunches in the beam is set by \vn{n_bunch}. If reading the distribution from a file,
if \vn{%n_bunch} is zero, the number of bunches created is the number of defined in the file and if
\vn{%n_bunch} is not zero, the number created is \vn{%n_bunch}. It is an error if \vn{%n_bunch} is
greater than the number of bunches defined in the file. If not reading from a file, if \vn{%n_bunch}
is zero, one bunch is created.
%
\item[\%n_particle] \Newline
Number of particles generated when the \vn{%distribution_type} is \vn{"RAN_GAUSS"}. Ignored for
other distribution types. When reading the distribution from a file, if \vn{%n_particle} is zero,
the number of particles in a bunch will be the number of particles defined in the file.  If
\vn{%n_particle} is non-zero when reading from a file, the number of particles in a bunch will be
\vn{%n_particle}. It is an error if \vn{%n_particle} is non-zero and the number of particles defined
in the file is less than \vn{%n_particle}.
%
\item[\%random_engine] \Newline
This component sets the algorithm to use in generating a uniform distribution
of random numbers in the interval [0, 1]. \vn{"pseudo"} is a pseudo random
number generator and "quasi" is a quasi random generator. "quasi random" is
a misnomer in that the distribution generated is fairly uniform.
%
\item[\%random_gauss_converter, \%random_sigma_cutoff] \Newline
To generate Gaussian random numbers, a conversion algorithm from the
flat distribution generated according to \vn{%random_engine} is
needed.  \vn{%random_gauss_converter} selects the algorithm. The
\vn{"exact"} conversion uses an exact conversion. The \vn{"quick"}
method is somewhat faster than the \vn{"exact"} method but not as accurate.
With either conversion method, if \vn{%random_sigma_cutoff} is set to a positive number,
this limits the maximum sigma generated.
%
\item[\%renorm_center, \%renorm_sigma] \Newline 
If set to True, these components will ensure that the actual beam center 
and sigmas will correspond to the input values. 
Otherwise, there will be fluctuations due to the finite number of 
particles generated.
%
\item[\%sig_pz, \%sig_z] \Newline
Longitudinal sigmas. \vn{%sig_pz} is the fractional energy spread dE/E.  This, along with
\vn{%dPz_dz} determine the longitudinal profile.

When simulating a ring, if any of these parameters is set negative, and if the \bmad based program
being run has enabled it, the value of the parameter will be set the value as calculated from the
lattice using synchrotron radiation integral formulas (\sref{s:synch.ints}).
%
\item[\%species] \Newline
Name of the species tracked. If not set then the default tracking particle type is used.
%
\item[\%spin] \Newline
Particle spin in Cartesian $(x, y, z)$ coordinates. Only used when not reading in particle positions
from a file. Also: When \vn{%use_particle_start} is set to True, and not reading in particle
positions form a file, the \vn{particle_start} spin values are used instead. See the description of
\vn{%use_particle_start} below for more details.
%
\item[\%use_particle_start] \Newline
If \vn{%use_particle_start} is set to \vn{False} (the default) the center of
the bunch is determined by the setting of the \vn{%center} component. If set to \vn{True}, the
center is determined by the setting of:
\begin{example}
  particle_start[x], particle_start[px],
  particle_start[y], particle_start[py],
  particle_start[z], particle_start[pz]
\end{example}
If not reading from from a particle position file, \vn{%use_particle_start} will affect the setting
of the spin. If \vn{%use_particle_start} is set to \vn{False}, the initial spin orientation is
determined by the setting of the \vn{%spin} component. If set to \vn{True}, the spin is determined
by:
\begin{example}
  particle_start[spin_x], particle_start[spin_y], particle_start[spin_z]
\end{example}
See \sref{s:beam.start} for details about the \vn{particle_start} structure. Using the
\vn{particle_start} structure allows setting center and spin values in the lattice file rather than
the setting of \vn{%center} and \vn{%spin} in the \vn{beam_init_struct}. In this case, \vn{%center}
is a dependent parameter and will be set to the value of \vn{particle_start}.
%
\item[\%use_t_coords, \%use_z_as_t] \Newline
The problem with handling particle distributions at low energies is that when a particle's velocity
is zero, the phase space $z$-coordinate is zero. What is needed here is to generate distributions in
time and or in longitudinal $s$-position. To do this, two switches, \vn{use_to_coords} and
\vn{use_z_as_t}, are used.

If \vn{use_t_coords} is True (default is False), the values in the distributions are taken as
describing particles using ``time coordinates'' (\sref{s:time.phase.space}).

The \vn{use_z_as_t} parameter is only used if \vn{use_t_coords} is set to True. When
\vn{use_t_coords} is True, if \vn{use_z_as_t} is also True, the value of the time (in seconds)
assigned to a particle will be set equal to the \vn{z}-coordinate (in meters) and a new
\vn{z}-coordinate will be calculated based upon \Eq{zbctt}.  This is useful for describing the
situation, for example, where particles may originate at a cathode at the same $s$-position, but
different times.  If \vn{use_z_as_t} is False (the default), the $z$ coordinate is taken to describe
the $s$-coordinate.  This is useful for modeling particles that have the same time but different $s$
positions. In this case, particles are ``born'' inside the lattice element at the location of the
particle's $s$-coordinate. To properly track the particles, the bunch will need to be tracked
through the element where they are born with a tracking method that can handle inside
particles. Currently the only such tracking method is \vn{time_runge_kutta}.  When the particles get
to the end of the element, the particle positions are converted to standard $s$-coordinates.

Since HDF5 beam files store complete information about a particle, \vn{use_to_coords} and
\vn{use_z_as_t} are not needed and are ignored.

\end{description}

%-----------------------------------------------------------------
\section{File Based Beam Initialization}
\label{s:beam.init.file}
\index{beam initialization parameters|hyperbf}

A beam initialization file specifies the coordinates of all the particles in a beam. If a \bmad
based program uses a \vn{beam_init_struct} (\sref{s:beam.init.file}) for inputting initialization
parameters, the file name for file based beam initialization can be set using the \vn{%position_file}
component of the structure. Also the bunch charge, bunch number, and number of particles per bunch
can be set in the \vn{beam_init_struct}. Additionally, the bunch centroid can be offset by setting
\vn{beam_init%center} and \vn{beam_init%center_jitter}.

There are two formats for the beam initialization file: ASCII and binary. The binary file format for
beam position storage is based on the \vn{HDF5} standard. More information is in
chapter~\sref{c:hdf5}.


The new ASCII format describes a particle bunch with a \vn{header} section followed by a \vn{table}
of particle parameters. Multiple bunches that comprise a beam can be specified by multiple header
section / particle parameter table pairs, one pair for each bunch. An example header section:
\begin{example}
\# The header field lines all start with a pound "\#" sign.
\# Any line in the header field that does not start with a recognized parameter or
\# does not have an equal sign is ignored.
\# my_param = 1.23    ! Custom parameters can be defined and will be ignored by Bmad.
\# species = proton   ! This parameter will be read by Bmad.
\# spin = 1, 0, 0     !   and this one too.
\# The last line in the header field starts with "\#!" and defines the table columns.
\#!   index     x         px       y     py  ... etc...
\end{example}
The header section lines all start with a pound ``\#'' sign. The last line in the header section
must start with ``\#!'' and this line defines the particle parameter table columns. With the
exception of the last line, all header lines will be ignored except ones that begin with a
recognized parameter followed by an equal sign. Recognized parameters are the components of the
\vn{coord_struct}, as documented in \sref{s:coord.struct}, that describes individual particles. In
addition, the following parameters are recognized:
\begin{example}
  charge_tot      ! Total bunch charge (including dead particles).
  s_position      ! Longitudinal position of bunch. Can be used in place of "s".
  time            ! Time particles of bunch are at. Can be used in place of "t".
\end{example}
These parameters, if present, will be used to set the corresponding particle parameter in all the
particles of the corresponding bunch. Possible \vn{location} and \vn{state} parameter settings are
documented in \sref{s:coord.struct}. The string equivalent to any setting is obtained by removing
the trailing ``\$'' from the variable name. For example, the variable \vn{alive\$} which is a
possible \vn{state} setting becomes the string \vn{alive}. Note: \vn{charge_tot} (total bunch
charge) will superceed \vn{charge} (the charge per particle).

The particle parameter table follows the header section. Each row gives the parameters for one
particle. Not all particle parameters must be specified. If a particular parameter is not present,
its default value will be used or, if present, the value given in the header section. Column order
is irrelavent and what determines what particle parameter is associated with a given column is the
last line of the header section which starts with the characters \vn{\#!}. Except for vector
parameters, column names correspond to \vn{coord_struct} parameters the same as in the header
section. For parameters that are vectors, the mapping from column name to parameter is:
\begin{example}
  Column Name             Corresponding coord_struct components
  ----------------------  -----------------------------------------
  x, px, y, py, z, pz     vec(1), ..., vec(6)
  spin_x, spin_y, spin_z  spin(1), spin(2), spin(3)
  field_x, field_y        field(1), field(2)
  phase_x, phase_y        phase(1), phase(2)
\end{example}
Additionally, there can be (but is not required) an \vn{index} column giving the particle index in
the array of particles of a bunch.  This column is ignored so, for example, values in the first line
of the table will always be used to set the particle with index 1 independent of the
value given in the index column. This behavior is implemented so that a beam file can be edited to add
or remove particles without worrying about reindexing.

Note: Example beam files (both ASCII and HDF5) can be generated using the program in the Bmad
Distribution \vn{code_examples/beam_track_example}.

Note: Beam files can be converted between ASCII and HDF5 binary using the program in the Bmad
Distribution \vn{util_programs/beam_file_translate_format}.

%-----------------------------------------------------------------
\subsection{Old Beam ASCII Format}

There is an old ASCII format that is still accepted by \bmad but is deprecated and should be
avoided. The format is:
\begin{example}
  <ix_ele>         ! Lattice element index. This is ignored.
  <n_bunch>        ! Number of bunches.
  <n_particle>     ! Number of particles per bunch to use
  [bunch loop: ib = 1 to n_bunch]
    BEGIN_BUNCH    ! Marker to mark the beginning of a bunch specification block.
    <species_name> ! Species of particle
    <charge_tot>   ! Total charge of bunch (alive + dead). 0 => Use <macro_charge>.
    <z_center>     ! z position at center of bunch.
    <t_center>     ! t position at center of bunch.
    [particle loop: Stop when END_BUNCH marker found]
      <x> <px> <y> <py> <z> <pz> <macro_charge> <state> <spin_x> <spin_y> <spin_z> 
    [end particle loop]
    END_BUNCH      ! Marker to mark the end of the bunch specification block
  [end bunch loop]
\end{example}
Example:
\begin{example}
  0       ! ix_ele
  1       ! n_bunch
  25000   ! n_particle
  BEGIN_BUNCH
    POSITRON
    3.2E-9   ! charge_tot
    0.0      ! z_center
    0.0      ! t_center
   -6.5E-3  9.6E-3 -1.9E-2  8.8E-3  2.2E-2 -2.4E-2  1.2E-13  1 1.0 0.0 0.0
    8.5E-3  5.5E-3  4.0E-2 -1.9E-2 -4.9E-3  2.1E-2  1.2E-13  1 1.0 0.0 0.0
    1.1E-2 -1.9E-2 -2.5E-2  1.0E-2 -1.8E-2 -7.1E-3  1.2E-13  1 1.0 0.0 0.0
   -3.4E-2 -2.7E-3 -4.1E-3  1.3E-2  1.3E-2  1.0E-2  1.2E-13  1 1.0 0.0 0.0
    6.8E-3 -4.5E-3  2.5E-3  1.4E-2 -2.3E-3  7.3E-2  1.2E-13  1 1.0 0.0 0.0
    1.2E-2 -9.8E-3  1.7E-3  6.4E-3 -9.8E-3 -7.2E-2  1.2E-13  1 1.0 0.0 0.0
    1.1E-2 -3.5E-4  1.2E-2  1.8E-2  5.4E-3  1.4E-2  1.2E-13  1 1.0 0.0 0.0
       ... etc. ...
  END_BUNCH
\end{example}

The first line of the file gives \vn{<ix_ele>}, the index of the lattice element at which the
distribution was created. This is ignored when the file is Read. 

The second line gives \vn{<n_bunch>}, the number of bunches. This can be overridden by a non-zero
setting of \vn{beam_init%n_bunch}.

The third line gives \vn{<n_particle>} the number of particles in a bunch. The actual number rows
specifying particle coordinates may be more then \vn{<n_particle>}. In this case, particles will be
discarded so that the beam has \vn{<n_particle>} particles.  The setting of \vn{<n_particle>} can be
overridden by a non-zero setting of \vn{beam_init%n_particle}.

After this, there are \vn{<n_bunch>} blocks of data, one for each bunch. Each one of these blocks
starts with a \vn{BEGIN_BUNCH} line to mark the beginning of the block and ends with a
\vn{END_BUNCH} marker line. In between, the first four lines give the \vn{<species_name>} name,
\vn{<charge_tot>}, \vn{<z_center>}, and \vn{<t_center>} values. The \vn{<species_name>} name may
be one of:
\begin{example}
  positron  ! default
  electron
  proton
  antiproton
  muon
  antimuon
  photon
\end{example}

The lines following the \vn{<t_center>} line specify particle coordinates. One line for each
particle.  Only the first six numbers, which are the phase space coordinates, need to be specified
for each particle. If the \vn{<macro_charge>} column is not present, or is zero, it defaults to
\vn{<charge_tot>/<n_particle>}.

The \vn{<state>} parameter indicates whether a particle is alive or dead. Values are
\begin{example}
  1     ! Alive
  2-7   ! Dead
  8     ! Pre-born
\end{example}
The \vn{pre-born} state indicates that the particle is waiting to be emitted from the cathode of 
an electron gun (\sref{s:sc.fft}).

The particle spin is specified by $x$, $y$ and $z$ components.

Each particle has an associated \vn{<macro_charge>}. If \vn{<charge_tot>} is set to a non-zero
value, the charge of all the particles will be scaled by a factor to make the total macro charge
equal to \vn{<charge_tot>}. The macro charge is ignored in tracking. The charge of the particle
used in tracking is the charge as calculated for the particle species. On the other hand, the macro
charge is used to calculate such things as the total charge in a particular region or the field
produced by a particle. That is, the macro charge acts as a weighting factor for a particle when the
particle's field or the particle's effect on other particles is calculated.

\index{change}\index{beam_init}
When the particle coordinates are read in the centroid will be shifted by the setting of
\vn{beam_init%center} (unless \vn{beam_init%use_particle_start} is set True) and
\vn{beam_init%center_jitter}.

