\begin{thebibliography}{XXXXXXX99}

\bibitem[Abell06]{b:rf.abell}
Dan Abell, ``Numerical computation of high-order transfer maps for rf cavities'',
Phys. Rev. ST Accel. Beams, {\bf 9}, pp. 052001, (2006).

\bibitem[Akima70]{b:akima}
H. Akima, 
``A New Method of Interpolation and Smooth Curve Fitting Based on Local Procedures,'' 
J.\ Assoc.\ Comp.\ Mach., {\bf 17} pp.~589-602 (1970).

\bibitem[Alianelli04]{b:mosaic}
L.~Alianeli, M.~Sanchez del Rio, R.~Fe1ici,
``I. A Monte Carlo algorithm for the simulation of Bragg scattering by imperfect crystals; 
and II. Application to mosaic copper,''
Proc.~SPIE 5536, Advances in Computational Methods for X-Ray and Neutron Optics, (2004).

\bibitem[Barber85]{b:barber85}
D.\ P.\ Barber, J.\ Kewisch, G.\ Ripken, R.\ Rossmanith, and R.\ Schmidt,
``A solenoid spin rotator for large electron storage rings,''
Particle Accelerators, {\bf 17} 243–262 (1985).

\bibitem[Barber99]{b:barber99}
D.~P.~Barber and G.~Ripken,
``Computer Algorithms and Spin Matching,''
{\it Handbook of Accelerator Physics and Engineering,}
1st Edition, 3rd printing, World Scientific Pub.\ Co.\ Inc.\ (1999).
Note: 1st edition has more detail than the 2nd edition.

\bibitem[Bater64]{b:batterman}
B.~Batterman, and H.~Cole,
``Dynamical Diffraction of X Rays by Perfect Crystals'',
Rev.\ Mod.\ Phys.,{\bf 36}, 3, pp.~ 681--717, (1964).

\bibitem[Bengt97]{b:bengtsson}
J.~Bengtsson,
``The Sextupole Scheme for the Swiss Light Source (SLS): An Analytic Approach,''
SLS Note 9/97, Paul Scherrer Institut, (1997).

\bibitem[Berz89]{b:berz}
M.~Berz, 
``Differential Algebraic Description of Beam Dynamics to Very High Orders,''
Particle Accelerators, Vol. 24, pp. 109-124, (1989).

\bibitem[Blas94]{b:blasdell}
R.~C.~Blasdell and A.~T.~Macrander, 
``Modifications to the 1989 SHADOW ray-tracing code for general asymmetric perfect-crystal optics,''
Nuc.\ Instr.\ \& Meth. A {\bf 347}, 320 (1994).

\bibitem[Bmad]{b:bmad.web}
The Bmad web site:
\hfill\break
\hspace*{0.3in} \url{https://www.classe.cornell.edu/bmad}

\bibitem[Brown77]{b:transport.appendix} 
K. L. Brown, F. Rothacker, D. C. Carey, and Ch. Iselin, 
``TRANSPORT Appendix,'' 
Fermilab, unpublished, (December 1977).

\bibitem[Chao79]{b:chao79}
Alexander W. Chao, 
``Evaluation of beam distribution parameters in an electron storage ring,''
Journal of Applied Physics {\bf 50}, 595 (1979).

\bibitem[Chao81]{b:chao.spin}
Alexander W. Chao, 
``Evaluation of Radiative Spin Polarization in an Electron Storage Ring,''
Nucl.\ Instrum.\ \& Meth. {\bf 180}, 29, (1981).

\bibitem[Chao93]{b:chao} 
Alexander Chao, 
{\em Physics of Collective Beam Instabilities in High Energy Accelerators}, 
Wiley, New York (1993). 

\bibitem[Corbett99]{b:corbett}
J. Corbett and Y. Nosochkov, ``Effect of Insertion Devices in SPEAR--3,''
Proc. 1999 Part.\ Acc.\ Conf., p.~238, (1999).

\bibitem[Decking00]{b:decking}
W. Decking, R. Brinkmann, 
``Space Charge Problems in the TESLA Damping Ring'',
Europ.\ Accel.\ Phys.\ Conf., Vienna, (2000).

\bibitem[Derby09]{b:derby}
Norman Derby and Stanislaw Olbert, 
``Cylindrical Magnets and Ideal Solenoids'',
arXiv:0909.3880v1 [physics.class-ph],
https://doi.org/10.48550/arXiv.0909.3880, (2009)


\bibitem[Duan15]{b:duan15}
Zhe Duan, Mei Bai, Desmond P. Barber, Qing Qin, 
``A Monte-Carlo simulation of the equilibrium beam polarization in ultra-high energy electron
(positron) storage rings,''
Nuc. Instrum. and methods in Phys. Research A {\bf 793}, pp 81-91, (2015).

\bibitem[Duff87]{b:leduff}
  J. Le Duff, \emph{Single and Multiple Touschek Effects}.
  Proc. CAS Berlin 1987,
  CERN 89-01, (1987).

\bibitem[Edwards73]{b:edteng}
D. A. Edwards and L. C. Teng, 
"Parametrization of Linear Coupled Motion in Periodic Systems",
IEEE Trans. Nucl. Sci. {\bf 20}, 3, (1973).

\bibitem[Elegant]{b:elegant}
The Elegant web site:
\hfill\break
\hspace*{0.3in} \url{https://ops.aps.anl.gov/elegant.html}

\bibitem[Forest02]{b:ptc}
\'E. Forest, F. Schmidt, E. McIntosh, 
{\it Introduction to the Polymorphic Tracking Code}, 
CERN-SL-2002-044 (AP), and KEK-Report 2002-3 (2002). 
Can be obtained at:
\hfill\break
\hspace*{0.3in}
\url{https://frs.web.cern.ch/frs/report/sl-2002-044.pdf}

\bibitem[Forest06]{b:geo.int}
\'Etienne Forest, `Geometric integration for particle accelerators,''
J. Phys. A: Math. Gen. {\bf 39} (2006) 5321–5377.

\bibitem[Forest88]{b:quad.fringe}
\'E. Forest, J. Milutinovic, 
``Leading Order Hard Edge Fringe Fields Effects Exact in ($1+\delta$) and 
Consistent with Maxwell's Equations for Rectilinear Magnets,''
Nuc. Instrum. and Methods in Phys. Research A {\bf 269}, pp 474-482, (1988).

\bibitem[Forest98]{b:forest}
\'E. Forest, {\em Beam Dynamics: A New Attitude and Framework},
Harwood Academic Publishers, Amsterdam (1998).

\bibitem[Grote96]{b:maduser}
H. Grote, F. C. Iselin, {\it The MAD Program User's Reference Manual},
Version 8.19, CERN/SL/90-13 (AP) (REV. 5) (1996). 
Can be obtained at:
\hfill\break
\hspace*{0.3in}
\url{https://mad.web.cern.ch/mad} 

\bibitem[Hartman93]{b:hartman}
S. C. Hartman and J. B. Rosenzweig, 
``Ponderomotive focusing in axisymmetric rf linacs'',
Phys. Rev. E {\bf 47}, 2031 (1993).

\bibitem[HDF5]{b:hdf5}
Hierarchical Data Format Version 5. Web page:
\hfill\break
\hspace*{0.3in} \url{https://www.hdfgroup.org/}

\bibitem[Healy86]{b:healy}
L. M. Healy, {\it Lie Algebraic Methods for Treating Lattice Parameter
Errors in Particle Accelerators}. Doctoral thesis, University of
Maryland, unpublished, (1986).

\bibitem[Helm73]{b:helm}
R. H. Helm, M. J. Lee, P. L. Morton, and M. Sands, 
``Evaluation of Synchrotron Radiation Integrals,''
IEEE Trans.~Nucl.~Sci. NS-20, 900 (1973).

\bibitem[Hoff06]{b:spin.hoff}
G.~Hoffstaetter, 
{\it High-Energy Polarized Proton Beams, A Modern View}, 
Springer. Springer Tracks in Modern Physics Vol~218, (2006).

\bibitem[Hwang15]{b:hwang}
Hwang and S.~Y.~Lee, 
``Dipole Fringe Field Thin Map for Compact Synchrotrons'',
Phys. Rev. ST Accel. Beams, {\bf 12}, 122401, (2015).

\bibitem[Iselin94]{b:madphysics}
F. C. Iselin, 
{\it The MAD program Physical Methods Manual}, 
unpublished, (1994). Can be obtained at: 
\hfill\break
\hspace*{0.3in}
\url{https://mad.home.cern.ch/mad}

\bibitem[Jackson76]{b:jackson}
J.~D.~Jackson,
``On understanding spin-flip synchrotron radiation and the transverse 
polarization of electrons in storage rings'',
Rev.\ Modern Phys., {\bf 48}, 3, 1976.

\bibitem[Jowett87]{b:jowett} 
J. M. Jowett, 
``Introductory Statistical Mechanics for Electron Storage Rings,'' 
AIP Conf. Proc. 153, Physics of Part.\ Acc.,
M. Month and M. Dienes Eds., pp.~864, (1987).

\bibitem[Kim11]{b:kim}
Hung Jin Kim,
``Symplectic map of crab cavity''
Unpublished, FERMILAB-TM-2523-APC.

\bibitem[Kohn95]{b:kohn}
V.~G.~Kohn, 
``On the Thcory of Reflectivitlby an X-Ray Multilaler Mirror''
physica status solidi (b), {\bf 187}, 61, (1995).

\bibitem[Lee99]{b:lee}
S.~Y.~Lee,
{\it Accelerator Physics},
World Scientific, Singapore, (1999).

\bibitem[Lynch90]{b:lynch}
Gerald Lynch and Orin Dahl,
``Approximations to Multiple Coulomb Scattering'',
Nucl. Instrum. Meth., B58, 6-10 (1991).

\bibitem[MAD-NG]{b:mad.ng}
Next Generation MAD program at:
\hfill\break
\hspace*{0.3in}
\url{https://github.com/MethodicalAcceleratorDesign/MAD}

\bibitem[McMill75]{b:mcmillan}
E. M. McMillan, 
``Multipoles in Cylindrical Coordinates,''
Nucl. Instrum. Meth. 127, 471 (1975).

\bibitem[Newton99]{b:newton}
D. Newton and A. Wolski, 
``Fast, Accurate Calculation of Dynamical Maps from
Magnetic Field Data Using Generalised Gradients,''
Proc. PAC09 (2009).

\bibitem[Ohmi94]{b:ohmi}
K.\,Ohmi, K.\,Hirata, and K.\,Oide,
``From the Beam-Envelope Matrix to Synchrotron-Radiation Integrals'',
Phys. Rev. E $\mathbf{49}$, 751 (1994).

\bibitem[OpenPMD]{b:openpmd}
openPMD: Open standard for particle-mesh data files. Documentation at:
\hfill\break
\hspace*{0.3in}
\url{https://github.com/openPMD/openPMD-standard}

\bibitem[Peralta12]{b:peralta}
Luis Peralta and Alina Louro, 
``AlfaMC: a fast alpha particle transport Monte Carlo code,''
 arXiv:1211.5960v3 [physics.comp-ph], 2012.

\bibitem[PGPLOT]{b:pgplot}
PGPLOT graphics library developed by Tim Pearson. Documentation at:
\hfill\break
\hspace*{0.3in}
\url{http://www.astro.caltech.edu/~tjp/pgplot/}

\bibitem[Piwin98]{b:piwinski}
Anton Piwinski, \emph{The Touschek Effect in Strong Focusing Storage Rings}.
DESY Report 98-179, 1998.

\bibitem[PLPLOT]{b:plplot}
PLPLOT graphics library developed primarily by Maurice J. LeBrun and Geoffrey Furnish. Documentation at:
\hfill\break
\hspace*{0.3in}
\url{http://plplot.sourceforge.net/}

\bibitem[Quat]{b:quat}
Wikipedia article, ``Quaternions and spatial rotation'',
\hfill\break
\hspace*{20pt} 
\url{https://en.wikipedia.org/wiki/Quaternions_and_spatial_rotation}.

\bibitem[Rauben91]{b:tol}
T. Raubenheimer,
``Tolerances to Limit the Vertical Emittance in Future Storage Rings'', 
Particle Accelerators, 1991, {\bf 36}, pp.75-119. 
SLAC-PUB-4937 Rev., (1991).

\bibitem[Rio98]{b:del.rio}
Manuel Sanchez del Rio, ``Ray tracing simulations for crystal optics,''
Proc. SPIE 3448, Crystal and Multilayer Optics, {\bf 230} (1998). 

\bibitem[Rosen94]{b:rosenzweig}
J. Rosenzweig and L. Serafini, ``Transverse Particle Motion in
Radio--Frequency Linear Accelerators,'' Phys Rev E, Vol. 49, p. 1599,
(1994).

\bibitem[Ruth87]{b:ruth} 
R. D. Ruth, ``Single-Particle Dynamics in
Circular Accelerators,'' in AIP Conference Proceedings {\bf 153}, {\em
Physics of Particle Accelerators}, pp.~152--235, M. Month and M. Dienes editors,
American Institute of Physics, New York (1987).

\bibitem[Ryne18]{b:opensc}
C. E. Mayes, R. D. Ryne, D. C. Sagan, 
``3D SPACE CHARGE IN BMAD'',
9\Th Int. Part. Accel. Conf., Vancouver, pp.~3428--3430, (2018).
Software available at: \url{https://github.com/RobertRyne/OpenSpaceCharge}.

\bibitem[SAD]{b:sad} 
D.~Zhou and K.~Oide, ``Maps Used in SAD'' (unpublished).
Also see:
\hfill\break
\hspace*{0.3in} \url{https://acc-physics.kek.jp/SAD/}

\bibitem[Sagan91]{b:beamion}
D. Sagan,
``Some aspects of the longitudinal motion of ions in electron storage rings,''
Nuc.\ Instr.\ \& Meth. A {\bf 307} 2, 171 (1991).


\bibitem[Sagan03]{b:wiggler}
D. Sagan, J. Crittenden, and D. Rubin,
``A Symplectic Model for Wigglers,'' Part.\ Acc.\ Conf. (2003).

\bibitem[Sagan99]{b:coupling}
D. Sagan and D. Rubin, 
``Linear Analysis of Coupled Lattices,''
Phys.\ Rev.\ ST Accel.\ Beams {\bf 2}, 074001 (1999).
\hfill\break
\hspace*{20pt} 
\url{https://link.aps.org/doi/10.1103/PhysRevSTAB.2.074001}

\bibitem[Sagan09]{b:csr}
D. Sagan, G. Hoffstaetter, C. Mayes, and U. Sae-Ueng,
``Extended one-dimensional method for coherent synchrotron radiation including shielding,''
Phys.\ Rev. Accel. \& Beams {\bf 12}, 040703 (2009).

\bibitem[Sagan17]{b:csr2}
D. Sagan and C. Mayes, 
``Coherent Synchrotron Radiation Simulations for Off-Axis Beams Using the Bmad Toolkit'',
Proc. Europ.\ Part.\ Accel.\ Conf. p. 2829 --- 31 (2006).

\bibitem[Sangwine]{b:biquat}
Stephen J.\ Sangwine, Todd A.\ Ell, Nicolas Le Bihan, 
``Fundamental Representations and Algebraic Properties of Biquaternions or Complexified Quaternions'',
Advances in Applied Clifford Algebras, 21, {\bf 21}, pp.~607, (2011).

\bibitem[Schoon11]{b:xraylib} 
T. Schoonjans et al. ``The xraylib library for X-ray-matter
interactions. Recent developments,'' Spectrochimica Acta Part B: Atomic
Spectroscopy {\bf 66}, pp. 776-784 (2011).
Code for xraylib is at: \hfill\break
\hspace*{0.3in}
\url{https://github.com/tschoonj/xraylib}

\bibitem[Silenko08]{b:silenko}
Alexander Silenko, 
``Equation of Spin Motion in Storage Rings in a Cylindrical Coordinate System,''
Phys.\ Rev.\ ST Accel.\ Beams, {\bf 9} pp. 034003, (2006).

\bibitem[Sommer18]{b:sommer}
Hannes Sommer, Igor Gilitschenski, Michael Bloesch, Stephan Weiss, Roland Siegwart, and Juan Nieto,
``Why and How to Avoid the Flipped Quaternion Multiplication'',
Aerospace {\bf 5}, 72, (2018), \hfill\break
\hspace*{0.3in}
\url{https://doi.org/10.3390/aerospace5030072}

\bibitem[Storn96]{b:de}
R.~Storn, and K.~V.~Price, ``Minimizing the real function of the
ICEC'96 contest by differential evolution'' IEEE conf. on Evolutionary
Computation, 842-844 (1996).

\bibitem[Sun10]{b:crab1}
Yi-Peng Sun, Ralph Assmann, Rogelio Tom\'as, and Frank Zimmermann,
``Crab Dispersion and its Impact on the CERN Large Hadron Collider Collimation'',
Phys. Rev. ST Accel. Beams, {\bf 13}, pp. 031001, (2010).

\bibitem[Talman87]{b:talman} R. Talman, ``Multiparticle Phenomena and
Landau Damping,'' in AIP Conf.\ Proc.  {\bf 153}, {\em Physics of
Particle Accelerators}, pp.~789--834, M. Month and M. Dienes editors,
American Institute of Physics, New York (1987).

\bibitem[Tao]{b:tao}
D. Sagan, J. Smith, {\it The Tao Manual}.
Can be obtained at: \hfill\break
\hspace*{0.3in}
\url{https://www.classe.cornell.edu/bmad/tao_entry_point.html}

\bibitem[Venturini98]{b:vent.map} 
M Venturini, D Abell, A Dragt, 
``Map Computation from Magnetic Field Data and Application to the LHC
High-Gradient Quadrupoles,'' 
Proc.\ 1998 International Comp.\ Accel.\ Phys.\ Conf.\ (ICAP '98),
{\bf 980914}, 184 (1998).

\bibitem[Venturini99]{b:gen.grad} 
M Venturini, A Dragt, 
``Accurate Computation of Transfer Maps from Magnetic Field Data'',
Nuc.\ Instr.\ \& Meth. A {\bf 427}, 387--392 (1999).

\bibitem[Wang12]{b:wang}
C.~Wang,
``Explicit formulas for 2nd-order driving terms due to sextupoles and
chromatic effects of quadrupoles,''
ANL/APS/LS-330, Argonne National Laboratory, (2012).

\bibitem[Wiede99]{b:wiedemann}
H. Wiedemann, {\em Particle Accelerator Physics}, Springer, New York, 3rd Edition (2007). 

\bibitem[Wolski06]{b:wolski.coupling}
A.~Wolski,  ``Alternative approach to general coupled linear optics,''
Phys. Rev. ST Accel. Beams 9, 024001 (2006).

\bibitem[Wyckoff65]{b:wyckoff}
R. W. G. Wyckoff, {\em Crystal Structures}, Interscience Publ. (1965).

\bibitem[Yazell09]{b:yazell}
Douglas J.~Yazell,
``Origins of the Unusual Space Shuttle Quaternion Definition'',
47th AIAA Aerospace Sciences Meeting, January, Orlando, Florida (2009).

\end{thebibliography}
