\section*{Introduction}
\pdfbookmark[1]{Introduction}{Intro}

As a consequence of \bmad being a software library, this manual serves two masters: The
programmer who wants to develop applications and needs to know about the inner workings of
\bmad, and the user who simply needs to know about the \bmad standard input format and
about the physics behind the various calculations that \bmad performs.

\index{MAD|hyperbf}
To this end, this manual is divided into three parts. The first two
parts are for both the user and programmer while the third part is
meant just for programmers. 
  \begin{description}
  \item[Part~I] \Newline
Part~I discusses the \bmad lattice input standard. The \bmad lattice input standard was
developed using the \mad\cite{b:maduser,b:madphysics}. lattice input standard as a
starting point but, as \bmad evolved, \bmad's syntax has evolved with it.
  \item[Part~II] \Newline
part~II gives the conventions used by \bmad --- coordinate systems, magnetic field
expansions, etc. --- along with some of the physics behind the calculations. By necessity,
the physics documentation is brief and the reader is assumed to be familiar with high
energy accelerator physics formalism.
  \item[Part~III] \Newline
Part~III gives the nitty--gritty details of the \bmad
subroutines and the structures upon which they are based.
\end{description}

\index{Bmad!information}
More information, including the most up--to--date version of this manual, can be found at
the \bmad web site\cite{b:bmad.web}.  Errors and omissions are a fact of life for any
reference work and comments from you, dear reader, are therefore most welcome. Please send
any missives (or chocolates, or any other kind of sustenance) to:
\begin{example}
  David Sagan <dcs16@cornell.edu>
\end{example}
\index{Bmad!error reporting}

The \bmad manual is organized as reference guide and so does not do a good job of instructing the
beginner as to how to use \bmad. For that there is an introduction and tutorial on \bmad and \tao
(\sref{s:tao.intro}) concepts that can be downloaded from the \bmad web page. Go to either the \bmad or
\tao manual pages and there will be a link for the tutorial.

It is my pleasure to express appreciation to people who have contributed to this effort, and without
whom, \bmad would only be a shadow of what it is today: To David Rubin and Georg Hoffstaetter for
their support all these years, to \'Etienne Forest (aka Patrice Nishikawa) for use of his remarkable
PTC/FPP library (not to mention his patience in explaining everything to me), to Desmond Barber for
very useful discussions on how to simulate spin, to Jonathan Laster, Mark Palmer, Matt Rendina, and
Attilio De~Falco for all their work maintaining the build system and for porting \bmad to different
platforms, to Frank Schmidt and CERN for permission to use the \mad tracking code. To Hans Grote and
CERN for granting permission to adapt figures from the \mad manual for use in this one, to Martin
Berz for his DA package, and to Dan Abell, Jacob Asimow, Ivan Bazarov, Moritz Beckmann, Scott Berg,
Oleksii Beznosov, Kevin Brown, Joel Brock, Sarah Buchan, Avishek Chatterjee, Jing Yee Chee, Christie
Chiu, Joseph Choi, Robert Cope, Jim Crittenden, Laurent Deniau, Bhawin Dhital, Gerry Dugan, Michael
Ehrlichman, Jim Ellison, Ken Finkelstein, Mike Forster, Thomas Gl{\"a}{\ss}le, Juan Pablo
Gonzalez-Aguilera, Sam Grant, Colwyn Gulliford, Eiad Hamwi, Klaus Heinemann, Richard Helms, Lucy
Lin, Henry Lovelace III, Chris Mayes, Vasiliy Morozov, Karthik Narayan, Katsunobu Oide, Tia Plautz,
Matt Randazzo, Robert Ryne, Michael Saelim, Jim Shanks, Matthew Signorelli, Hugo Slepicka, Jeff
Smith, Jonathan Unger, Jeremy Urban, Ningdong Wang, Suntao Wang, Mark Woodley, and Demin Zhou for
their help.

