\chapter{Normal Modes: Twiss Parameters, Coupling, Emittances, Etc.}
\label{c:normal.modes}

%-----------------------------------------------------------------------------
\section{Components in the Ele\_struct}
\label{s:twiss.ele}

\index{ele_struct!\%gamma_c}
\index{ele_struct!\%c_mat}
\index{ele_struct!\%a}\index{ele_struct!\%b}\index{ele_struct!\%z}
The \vn{ele_struct} (\sref{c:ele.struct}) has a number of components that hold information on the
Twiss parameters, dispersion, and coupling at the exit end of the element. The Twiss parameters of
the three normal modes (\sref{s:coupling}) are contained in the \vn{ele%a}, \vn{ele%b}, and
\vn{ele%z} components which are of type \vn{twiss_struct}:
\index{twiss_struct}
\begin{example}
  type twiss_struct
    real(rp) beta         ! Twiss Beta function
    real(rp) alpha        ! Twiss Alpha function
    real(rp) gamma        ! Twiss gamma function
    real(rp) phi          ! Normal mode Phase advance
    real(rp) eta          ! Normal mode dispersion
    real(rp) etap         ! Normal mode momentum dispersion.
    real(rp) deta_ds      ! Dispersion derivative
    real(rp) sigma        ! Normal mode beam size
    real(rp) sigma_p      ! Normal mode beam size derivative
    real(rp) emit         ! Geometric emittance
    real(rp) norm_emit    ! Energy normalized emittance (= \(\beta \, \gamma \, \epsilon\))
  end type 
\end{example}
\index{xy_disp_struct}
The projected horizontal and vertical dispersions in an \vn{ele_struct} are contained in the
\vn{ele%x} and \vn{ele%y} components. These components are of type \vn{xy_disp_struct}:
\index{xy_disp_struct}
\begin{example}
  type xy_disp_struct
    real(rp) eta     ! Projected dispersion 
    real(rp) etap    ! Projected momentum dispersion
    real(rp) deta_ds ! Projected dispersion derivative \(d\eta_x/ds\) or \(d\eta_y/ds\).
  end type 
\end{example}
Section~\sref{s:dispersion} discussed the relationship between \vn{etap} and \vn{deta_ds}.

\index{ele_struct!\%emit}\index{ele_struct!\%norm_emit}
\index{ele_struct!\%sigma}\index{ele_struct!\%sigma_p}
The components \vn{ele%emit}, \vn{ele%norm_emit}, \vn{ele%sigma}, \vn{ele%sigma_p} are not set by
the standard \bmad routines and are present for use by any program.

\index{ele_struct!\%c}\index{ele_struct!\%gamma_c}
The relationship between the projected and normal mode dispersions are given by \Eq{avx}. The 2x2
coupling matrix $\bfC$ (\Eq{vgicc1}) is stored in the \vn{ele%c_mat(2,2)} component of the
\vn{ele_struct} and the $\gamma$ factor of \Eq{vgicc1} is stored in the \vn{ele%gamma_c}
component. There are several routines to manipulate the coupling factors. For example:
\begin{example}
  \Hyperref{r:c.to.cbar}{c_to_cbar}(ele, cbar_mat)             ! Form Cbar(2,2) matrix
  \Hyperref{r:make.v.mats}{make_v_mats}(ele, v_mat, v_inv_mat)   ! Form V coupling matrices.
\end{example}
See \sref{r:mat} for a complete listing of such routines.

Since the normal mode and projected dispersions are related, when one is changed within a program
the appropriate change must be made to the other. To make sure everything is consistent, the
\Hyperref{r:set.flags.for.changed.attribute}{set_flags_for_changed_attribute} routine can be
used. Example:
\begin{example}
  type (lat_struct), target :: lat
  real(rp), pointer :: attrib_ptr
  ...
  attrib_ptr => lat%ele(ix_ele)%value(k1$) ! Point to some attribute.
  attrib_ptr = value                       ! Change the value.
  call set_flags_for_changed_attribute (lat%ele(ix_ele), attrib_ptr)
\end{example}

The \vn{%mode_flip} logical component of an \vn{ele_struct} indicates whether the $a$ and $b$ normal
modes have been flipped relative to the beginning of the lattice. See Sagan and
Rubin\cite{b:coupling} for a discussion of this. The convention adopted by \bmad is that the \vn{%a}
component of all the elements in a lattice will all correspond to the same physical normal
mode. Similarly, the \vn{%b} component of all the elements will all correspond to some (other)
physical normal mode.  That is, at an element where there is a mode flip (with \vn{%mode_flip} set
to True), the \vn{%a} component actually corresponds to the $\bfB$ matrix element in \Eq{ua00b} and
vice versa. The advantage of this convention is that calculations of mode properties (for example
the emittance), can ignore whether the modes are flipped or not.

The normal mode analysis of Sagan and Rubin, while it has the benefit of simplicity, is strictly
only applicable to lattices where the RF cavities are turned off.  The full 6-dimensional analysis
is summarized by Wolski\cite{b:wolski.coupling}.  The
\Hyperref{r:normal.mode3.calc}{normal_mode3_calc} routine perform the full analysis. The results are
put in the \vn{%mode3} component of the \vn{ele_struct} which is of type \vn{mode3_struct}:
\index{ele_struct!\%mode3}
\index{mode3_struct}
\begin{example}
  type mode3_struct
    real(rp) v(6,6)
    type (twiss_struct) a, b, c
    type (twiss_struct) x, y
  end type
\end{example}
The 6-dimensional \vn{mode3%v(6,6)} component is the analog of the 4-dimensional
$\bfV$ matrix appearing in \Eq{tvuv}.

%-----------------------------------------------------------------------------
\section{Tune and Twiss Parameter Calculations}
\label{s:twiss}
\index{twiss parameters}
\index{twiss parameters!calculation}
\index{tune!calculation}

A calculation of the Twiss parameters starts with the Twiss parameters at the beginning of the
lattice. For linear machines, these Twiss parameters are generally set in the input lattice file
(\sref{s:beginning}). For circular machines, the routine \Hyperref{r:twiss.at.start}{twiss_at_start}
may be used (\sref{s:beginning})
\begin{example}
  type (lat_struct) lat
  ...
  if (lat%param%geometry == closed$) call twiss_at_start(lat)
\end{example}
In either case, the initial Twiss parameters are placed in \vn{lat%ele(0)}.  The tune is placed in
the variables \vn{lat%a%tune} and \vn{lat%b%tune}.

To propagate the Twiss, coupling and dispersion parameters from the start of the lattice to the end,
the routine, \Hyperref{r:twiss.propagate.all}{twiss_propagate_all} can be used. This routine works
by repeated calls to \Hyperref{r:twiss.propagate1}{twiss_propagate1} which does a single propagation
from one element to another. The Twiss propagation depends upon the transfer matrices having already
computed (\sref{c:tracking}).  \vn{twiss_propagate_all} also computes the Twiss parameters for all
the lattice branches.

Before any Twiss parameters can be calculated, the transfer matrices stored in the lattice elements
must be computed.  \Hyperref{r:bmad.parser}{bmad_parser} does this automatically about the zero
orbit. If, to see nonlinear effects, a different orbit needs to be used for the reference, The
routine \Hyperref{r:twiss.and.track}{twiss_and_track} can be used. For example:
\begin{example}
  type (lat_struct) lat
  type (coord_struct), allocatable :: orbit(:)
  call bmad_parser ('my_lattice', lat)
  call twiss_and_track (lat, orb, ok)
\end{example}

Once the starting Twiss parameters are set, \Hyperref{r:twiss.propagate.all}{twiss_propagate_all}
can be used to propagate the Twiss parameters to the rest of the elements
\begin{example}
\end{example}

The routine \Hyperref{r:twiss.and.track.at.s}{twiss_and_track_at_s} can be used to calculate the
Twiss parameters at any given longitudinal location. Alternatively, to propagate the Twiss
parameters partially through a given element use the routine
\Hyperref{r:twiss.and.track.intra.ele}{twiss_and_track_intra_ele}.

%-----------------------------------------------------------------------------
\section{Tune Setting}
\label{s:tune.set}
\index{tune!setting}

The routine \Hyperref{r:set.tune}{set_tune} can be used
to set the transverse tunes:
\begin{example}
  set_tune (phi_a_set, phi_b_set, dk1, lat, orb_, ok)
\end{example}
\vn{set_tune} varies quadrupole strengths until the desired tunes are achieved. As
input,\vn{set_tune} takes an argument \vn{dk1(:)} which is an array that specifies the relative
change to be make to the quadrupoles in the lattice.

To set the longitudinal (synchrotron) tune, the routine \Hyperref{r:set.z.tune}{set_z_tune} can be
used.  \Hyperref{r:set.z.tune}{set_z_tune} works by varying rf cavity voltages until the desired
tune is achieved.

%-----------------------------------------------------------------------------
\section{Emittances \& Radiation Integrals}
\label{s:emit}

See Section~\sref{s:synch.ints} for details on the radiation integral formulas.

The routine \Hyperref{r:radiation.integrals}{radiation_integrals} is used to calculate the 
normal mode emittances along with the radiation integrals:
\begin{example}
  type (lat_struct) lat
  type (normal_modes_struct) modes
  type (rad_int_all_ele_struct) ele_rad_int
  ...
  call radiation_integrals (lat, orbit, modes, rad_int_by_ele = ele_rad_int)
\end{example}
The \vn{modes} argument, which is of type \vn{normal_modes_struct}, holds the radiation integrals
integrated over the entire lattice.
\begin{example}
  type normal_modes_struct
    real(rp) synch_int(0:3) ! Synchrotron integrals I0, I1, I2, and I3
    real(rp) sigE_E         ! SigmaE/E
    real(rp) sig_z          ! Sigma_Z
    real(rp) e_loss         ! Energy loss / turn (eV)
    real(rp) rf_voltage     ! Total rfcavity voltage (eV)
    real(rp) pz_aperture    ! pz aperture limit
    type (anormal_mode_struct)  a, b, z
    type (linac_normal_mode_struct) lin
  end type
\end{example}
In particular, the \vn{%a}, \vn{%b}, and \vn{%z} components, which are of type
\vn{anormal_mode_struct} hold the emittance values:
\begin{example}
  type anormal_mode_struct
    real(rp) emittance        ! Beam emittance
    real(rp) synch_int(4:6)   ! Synchrotron integrals
    real(rp) j_damp           ! damping partition number
    real(rp) alpha_damp       ! damping per turn
    real(rp) chrom            ! Chromaticity
    real(rp) tune             ! "Fractional" tune in radians
  end type
\end{example}

The \vn{ele_rad_int} argument, which is is of type \vn{rad_int_all_ele_struct}, holds the radiation
integrals on an element-by-element basis.
\begin{example}
  type rad_int_all_ele_struct
    type (rad_int1_struct), allocatable :: ele(:) ! Array is indexed from 0
  end type
\end{example}

%-----------------------------------------------------------------------------
\section{Chromaticity Calculation}
\label{s:chrom}

\index{chromaticity}
For a circular lattice, \Hyperref{r:chrom.calc}{chrom_calc} calculates the chromaticity by
calculating the tune change with change in beam energy.

\Hyperref{r:chrom.tune}{chrom_tune} sets the chromaticity by varying the sextupoles. This is a very
simple routine that simply divides the sextupoles into two families based upon the local beta
functions at the sextupoles.



