\section*{Overview}
\index{Bmad|hyperbf}
\pdfbookmark[1]{Overview}{Overview}

\bmad (Otherwise known as ``Baby MAD" or ``Better MAD" or just plain ``Be MAD!") is a subroutine
library for charged--particle and X-Ray simulations in accelerators and storage rings. \bmad has
been developed at the Cornell Laboratory for Accelerator-based ScienceS and Education (CLASSE) and
has been in use since 1996.

Prior to the development of \bmad, simulation programs at Cornell were written almost from scratch
to perform calculations that were beyond the capability of existing, generally available software.
This practice was inefficient, leading to much duplication of effort. Since the development of
simulation programs was time consuming, needed calculations where not being done. As a response, the
\bmad subroutine library, using an object oriented approach and written in Fortran 2008, were
developed.  The aim of the \bmad project was to:
\begin{Itemize}
\item Cut down on the time needed to develop programs.
\item Cut down on programming errors.
\item Provide a simple mechanism for lattice function calculations
from within control system programs.
\item Provide a flexible and powerful lattice input format.
\item Standardize sharing of lattice information between 
programs.
\end{Itemize}

\bmad can be used to study both single and multi--particle beam dynamics as well as X-rays.  Over
the years, \bmad modules have been developed for simulating a wide variety of phenomena including
intra beam scattering (IBS), coherent synchrotron radiation (CSR), Wakefields, Touschek scattering,
higher order mode (HOM) resonances, etc., etc.  \bmad has various tracking algorithms including
Runge--Kutta and symplectic (Lie algebraic) integration. Wakefields, and radiation excitation and
damping can be simulated. \bmad has routines for calculating transfer matrices, emittances, Twiss
parameters, dispersion, coupling, etc. The elements that \bmad knows about include quadrupoles, RF
cavities (both storage ring and LINAC accelerating types), solenoids, dipole bends, Bragg crystals
etc.  In addition, elements can be defined to control the attributes of other elements. This can be
used to simulate the ``girder'' which physically support components in the accelerator or to easily
simulate the action of control room ``knobs'' that gang together, say, the current going through a
set of quadrupoles.

To be able to extend \bmad easily, \bmad has been developed in a modular, object oriented, fashion
to maximize flexibility. As just one example, each individual element can be assigned a particular
tracking method in order to maximize speed or accuracy and the tracking methods can be assigned via
the lattice file or at run time in a program.

\vfill
\break
