\chapter{Parameter Structures}
\label{c:param.structs}

%-----------------------------------------------------------------
\section{What is a Structure?}
\label{s:struct}
\index{structure|hyperbf}

A ``structure'' is a collection of parameters.  \bmad has various structures which can be used for
various tasks. For example, the \vn{beam_init_struct} structure (\sref{s:beam.init.struct}) is used
to set parameters used to initialize particle beams.

A given program may give the user access to some of these structures so, in order to allow
intelligent parameter setting, this chapter gives an in-depth description of the most common ones.

Each structure has a \vn{``structure name''} (also called a \vn{``type name''}) which identifies the
list of parameters (also called ``components'') in the structure. Associated with a structure there
will be an \vn{``instance''} of this structure and this instance will have an \vn{``instance name''}
which is what the user uses to set parameters. It is possible to have multiple instances of a
structure. For example, in the situation where a program is simulating multiple particle beams,
there could be multiple \vn{beam_init_struct} (\sref{s:beam.init.struct}) instances with one for
each beam.

\bmad defines uses some structures to hold global parameters. That is, parameters that shared by all
code. Each of these structures has a single associated instance. These are:
\begin{center}
  \begin{tabular}{ll} \toprule
  Structure                 & Instance         \\
  \midrule
  bmad_common_stuct         & bmad_com         \\
  space_charge_common_stuct & space_charge_com \\
  \bottomrule
  \end{tabular}
\end{center}
All other structures will have instance names that are program specific. That is, see the program
documentation for the instance name(s) used.

For historical reasons, There are two syntaxes used for setting structure components. The syntax when
setting in a lattice file uses square brackets:
\begin{example}
  instance_name[parameter_name] = value
\end{example}
When setting a component in a program initialization file the syntax uses the percent ``\vn{%}'' character:
\begin{example}
  instance_name%parameter_name = value
\end{example}
Examples:
\begin{example}
  bmad_com[max_aperture_limit] = 10   ! Lattice file syntax.
  bmad_com%max_aperture_limit = 10    ! Program initialization file syntax.
\end{example}
this sets the \vn{max_aperture_limit} parameter of \vn{bmad_com} which is an instance name of the
\vn{bmad_common_struct}. Note: A program is free to set the instance name for any structure. This
should be documented in the program manual.

Note: Thought must be given to setting \vn{bmad_com} and other global parameters in a lattice file
(\sref{s:bmad.ptc.com}) since that will affect every program that uses the lattice.

%-----------------------------------------------------------------
\section{Bmad_Common_Struct}
\label{s:bmad.common}
\index{Bmad!general parameters|hyperbf}

The \vn{bmad_common_struct} structure contains a set of global parameters. There is only one global
instance (\sref{c:param.structs}) of this structure and this instance has the name
\vn{bmad_com}. The components of this structure along with the default values are:
\index{max_aperture_limit}\index{d_orb(6)}
\index{grad_loss_sr_wake}\index{default_ds_step}
\index{rel_tol_tracking}\index{abs_tol_tracking}
\index{rel_tol_adaptive_tracking}\index{abs_tol_adaptive_tracking}
\index{taylor_order}\index{default_integ_order}
\index{sr_wakes_on}\index{lr_wakes_on}\index{high_energy_space_charge_on}
\index{auto_bookkeeper}\index{csr_and_space_charge_on}
\index{spin_tracking_on}\index{radiation_damping_on}\index{spin_sokolov_ternov_flipping_on}
\index{radiation_fluctuations_on}\index{radiation_zero_average}\index{mp_threading_is_safe}
\index{absolute_time_tracking}
\index{bmad_common_struct|hyperbf}
\begin{example}
  type bmad_common_struct
    max_aperture_limit = 1e3            ! Max Aperture.
    d_orb(6)           = 1e-5           ! for the make_mat6_tracking routine.
    default_ds_step    = 0.2            ! Integration step size.  
    significant_length = 1e-10          ! meter 
    rel_tol_tracking = 1e-8             ! Closed orbit relative tolerance.
    abs_tol_tracking = 1e-11            ! Closed orbit absolute tolerance.
    rel_tol_adaptive_tracking = 1e-8    ! Runge-Kutta tracking relative tolerance.
    abs_tol_adaptive_tracking = 1e-10   ! Runge-Kutta tracking absolute tolerance.
    init_ds_adaptive_tracking = 1e-3    ! Initial step size.
    min_ds_adaptive_tracking = 0        ! Minimum step size to use.
    fatal_ds_adaptive_tracking = 1e-8   ! Threshold for loosing particles.
    autoscale_amp_abs_tol = 0.1_rp      ! Autoscale absolute amplitude tolerance (eV).
    autoscale_amp_rel_tol = 1e-6        ! Autoscale relative amplitude tolerance
    autoscale_phase_tol = 1e-5          ! Autoscale phase tolerance.
    electric_dipole_moment = 0          ! Particle's EDM. 
    synch_rad_scale = 1.0               ! Synch radiation kick scale. 1 => normal
    ptc_cut_factor = 0.006              ! Cut factor for PTC tracking
    sad_eps_scale = 5.0e-3              ! Used in sad_mult step length calc.
    sad_amp_max = 5.0e-2                ! Used in sad_mult step length calc.
    sad_n_div_max = 1000                ! Used in sad_mult step length calc.
    taylor_order = 3                    ! 3rd order is default
    default_integ_order = 2             ! PTC integration order
    ptc_max_fringe_order = 2            ! PTC max fringe order (2 => Quadrupole !).
    max_num_runge_kutta_step = 10000    ! Max num RK steps before particle is lost.
    rf_phase_below_transition_ref = F   ! Autoscale around phase phi0 = 0.5
    sr_wakes_on = T                     ! Short range wakefields?
    lr_wakes_on = T                     ! Long range wakefields
    ptc_use_orientation_patches = T     ! Ele orientation translated to PTC patches?
    auto_bookkeeper = T                 ! Automatic bookkeeping?
    high_energy_space_charge_on = F     ! High energy space charge calc toggle.
    csr_and_space_charge_on = F         ! CSR and space charge (separate from HE SC).
    spin_tracking_on = F                ! spin tracking?
    spin_sokolov_ternov_flipping_on = F ! Spin flipping during radiation emission?
    radiation_damping_on = F            ! Radiation damping toggle.
    radiation_fluctuations_on = F       ! Radiation fluctuations toggle.
    radiation_zero_average = F          ! Shift so that average radiation kick is zero?
    conserve_taylor_maps = T            ! Enable bookkeeper to set
                                        ! ele%taylor_map_includes_offsets = F?
    absolute_time_tracking = F          ! Set absolute or relative time tracking.
    absolute_time_ref_shift = T         ! Absolute time referenced to element ref time?
    aperture_limit_on = T               ! Use aperture limits in tracking.
    debug = F                           ! Used for code debugging.
  end type
\end{example}

Parameter description:
\begin{description}
\index{aperture_limit_on}
\item[\vn{abs_tol_adaptive_tracking}] \Newline
Absolute tolerance to use in adaptive tracking. This is used in \vn{runge-kutta} and
\vn{time_runge_kutta} tracking (\sref{s:integ}).
%
\item[\vn{abs_tol_tracking}] \Newline
Absolute tolerance to use in tracking. Specifically, Tolerance to use when finding the closed orbit.
%
\item[\vn{absolute_time_tracking}] \Newline
The \vn{absolute_time_tracking} switch\footnote
  {
An old, deprecated notation for this switch is \vn{parameter[absolute_time_tracking]}.
  }
sets whether the clock for the \vn{lcavity} and \vn{rfcavity} elements is tied to the reference
particle or to uses the absolute time (\sref{s:rf.time}). A value of \vn{False} (the default)
mandates relative time and a value of \vn{True} mandates absolute time. The exception is that for an
\vn{e_gun} element (\sref{s:e.gun}), absolute time tracking is always used in order to be able to
avoid problems with a zero reference momentum at the beginning of the element.
%
\item[\vn{absolute_time_ref_shift}] \Newline
When absolute time tracking is used (\sref{s:rf.time}), if \vn{absolute_time_ref_shift} is True (the
default), then the value of the time used to calculate RF phases and other time dependent parameters
is shifted by the reference time of the lattice element under consideration. If set to False, no
time shift is done. The advantage of \vn{absolute_time_ref_shift} set to True is that (at least on
the first turn of tracking) there is no phase shift between relative time and absolute time
tracking. The advantage of \vn{absolute_time_ref_shift} set to False is that when trying to compare
tracking in \bmad with tracking in programs that use absolute time tracking but do not implement a
reference shift (for example, the IMPACT and GPT programs), it is convenient not to have to worry
about the reference shift.
%
\item[\vn{aperture_limit_on]}] \Newline
Aperture limits may be set for elements in the lattice (\sref{s:limit}). Setting
\vn{aperture_limit_on} to \vn{False} will disable all set apertures. \vn{True} is the default.
%
\item[\vn{auto_bookkeeper}] \Newline
Toggles automatic or intelligent bookkeeping. See
section~\sref{s:lat.book} for more details.
%
\item[\vn{autoscale_amp_abs_tol}] \Newline
Used when \bmad autoscales (\sref{s:autoscale}) an elements field amplitude. This parameter sets the
absolute tolerance for the autoscale amplitude parameter.
%
\item[\vn{autoscale_amp_rel_tol}] \Newline
Used when \bmad autoscales (\sref{s:autoscale}) an elements field amplitude. This parameter sets the
relative tolerance for the autoscale amplitude parameter.
%
Used when \bmad autoscales (\sref{s:autoscale}) an elements AC phase. This parameter sets the
absolute tolerance for the autoscale parameter.
%
\item[\vn{autoscale_phase_tol}] \Newline
\item[\vn{init_ds_adaptive_tracking}] \Newline
Initial step to use for adaptive tracking. This is used in
\vn{runge-kutta} and \vn{time_runge_kutta} tracking (\sref{s:integ}).
%
\item[\vn{conserve_taylor_maps}] \Newline
Toggle to determine if the Taylor map for an element include any
element ``misalignments''.  See Section~\sref{s:mapoff} for more
details.
%
\item[\vn{csr_and_space_charge_on}] \Newline
Turn on or off the coherent synchrotron radiation and space charge calculations. (\sref{s:csr}).
The space charge calculation here is not to be confused with the high energy space charge
calculation (\sref{s:he.space.charge})
%
\item[\vn{d_orb}] \Newline 
Sets the orbit displacement used in the routine that calculates the transfer matrix through an
element via tracking. That is, when the \vn{mat6_calc_method} (\sref{s:xfer}) is set to
\vn{tracking}. \vn{d_orb} needs to be large enough to avoid significant round-off errors but not so
large that nonlinearities will affect the results. The default value is $10^{-5}$.
%
\item[\vn{debug}] \Newline
Used for communication between program units for debugging purposes.
%
\item[\vn{default_ds_step}] \Newline
Step size for tracking code \sref{c:methods} that uses a fixed step
size. For example, \vn{symp_lie_ptc} tracking.
%
\item[\vn{default_integ_order}] \Newline
Order of the integrator used by \'Etienne Forest's PTC code (\sref{s:libs}).
The order of the PTC integrator is like the order of a Newton-Cotes method.
Higher order means the error term involves a higher order derivative of the field.
%
\item[\vn{electric_dipole_moment}] \Newline
The electric dipole moment value used in tracking a particle's spin (\sref{s:spin.dyn}).
%
\item[\vn{fatal_ds_adaptive_tracking}] \Newline
This is used in \vn{runge-kutta} and \vn{time_runge_kutta} tracking
(\sref{s:integ}).  If the step size falls below the value set for
\vn{fatal_ds_adaptive_tracking}, a particle is considered lost.
This prevents a program from ``hanging'' due to taking a large number
of extremely small steps. The most common cause of small step size is
an ``unphysical'' magnetic or electric field.
%
\item[\vn{high_energy_space_charge_on}] \Newline
Toggle to turn on or off the ultra-relativistic space charge effect in particle tracking
(\sref{s:he.space.charge}). Computationally, this is separate from the lower energy space charge and
CSR calculation (\sref{s:csr}). Default is \vn{False}. Notice that including the high energy space
charge can be done on a branch-by-branch basis (\sref{s:beginning}).
%
\item[\vn{lr_wakes_on}] \Newline
Toggle for turning on or off long-range higher order mode wakefield effects.
%
\item[\vn{max_aperture_limit}] \Newline 
Sets the maximum amplitude a particle can have during tracking. If this amplitude is exceeded, the
particle is lost even if there is no element aperture set. Having a maximum aperture limit helps
prevent numerical overflow in the tracking calculations.
%
\item[\vn{max_num_runge_kutta_step}] \Newline 
The maximum number of steps to take through an element with \vn{runge_kutta} or
\vn{time_runge_kutta} tracking. The default value is 10,000. If the number of steps reaches this
value, the particle being tracked is marked as lost and a warning message is issued. Under
``normal'' circumstances, a particle will take far fewer steps to track through an element. If a
particle is not through an element after 10,000 steps, it generally indicates that there is a
problem with how the field is defined. That is, the field does not obey Maxwell's
Equations. Especially: discontinuities in the field can cause problems.
%
\item[\vn{min_ds_adaptive_tracking}] \Newline
This is used in \vn{runge-kutta} and \vn{time_runge_kutta} tracking (\sref{s:integ}). Minimum step
size to use for adaptive tracking. If To be useful, \vn{min_ds_adaptive_tracking} must be set
larger than the value of \vn{fatal_ds_adaptive_tracking}. In this case, particles are never lost
due to taking too small a step.
%
\item{ptc_use_orientation_patches} \Newline
With \bmad, there is no distinction whether an element's orientation attributes (\vn{offsets},
\vn{pitches}, or \vn{tilt} (\sref{s:offset})) is deliberate (part of the ``design'' of the machine)
or an error (a ``misalignment'').  With PTC this is not true. If the \vn{ptc_use_orientation_patches}
switch is set to True (the default), when a \bmad element is translated to PTC,
the element's orientation attributes are stored as patches. That is, ``design'' values.
If set to False, these parameters are stored as misalignments. This will generally not make any
difference to a calculation. The exception comes with PTC centric programs that vary machine
parameters.\footnote
  {
None of the programs that come bundled with \bmad (a \bmad \vn{Distribution}) are
PTC centric.
  }
%
\item{ptc_max_fringe_order} \Newline
Maximum order for computing fringe field effects in PTC. 
%
\item[\vn{rf_phase_below_transition_ref}] \Newline
Used when \bmad autoscales (\sref{s:autoscale}) an \vn{rfcavity} and when \bmad calculates the
reference time through a cavity (which affects calculation of phase space $z$ via \Eq{zbctt}).  If
True, the reference phase will be taken to be at \vn{phi0} = 0.5 which is appropriate for a ring
below transition. Default is False in which case autoscaling will be around the phase \vn{phi0} = 0.
%
\item[\vn{radiation_damping_on}] \Newline
Toggle to turn on or off effects due to radiation damping in particle tracking
(\sref{s:radiation}). The default is \vn{False}. Note: The standard \bmad emittance calculation,
which involves calculating synchrotron radiation integrals (\sref{s:synch.ints}) can be done without
a problem when \vn{radiation_damping_on} is set to False. However, since the closed orbit will be
affected by whether \vn{radiation_damping_on} is set or not, the calculated emittances will depend
upon the setting of \vn{radiation_damping_on}.
%
\item[\vn{radiation_fluctuations_on}] \Newline
Toggle to turn on or off effects due to radiation fluctuations in particle tracking
(\sref{s:radiation}).  The default is \vn{False}. Note: The standard \bmad emittance calculation,
which involves calculating synchrotron radiation integrals (\sref{s:synch.ints}) can be done without
a problem when \vn{radiation_damping_on} is set to False. And since the calculation of the closed orbit
ignores the fluctuating part of the radiation, the setting of \vn{radiation_damping_on}, unlike the
setting of \vn{radiation_damping_on}, will not affect the emittance calculation.
%
\item[\vn{radiation_zero_average}] \Newline
As discussed in Section~\sref{s:radiation}, it is sometimes convenient to shift the emitted radiation
spectrum so that the average energy emitted along the closed orbit is zero. This gets rid of the ``sawtooth''
effect. To shift the average emitted energy to zero, set \vn{radiation_zero_average} to \vn{True}. The
default is \vn{False}. Currently, the shifting of the spectrum only works for non PTC
dependent tracking. That is, the shifting is not applicable to tracking with Taylor maps and with
\vn{symp_lie_ptc} (\sref{s:tkm}) tracking.
%
\item[\vn{rel_tol_adaptive_tracking}] \Newline
Relative tolerance to use in adaptive tracking. This is used in \vn{runge_kutta} and
\vn{time_runge_kutta} tracking (\sref{s:integ}).
%
\item[\vn{rel_tol_tracking}] \Newline
Relative tolerance to use in tracking. Specifically, Tolerance to use when finding the closed orbit.
%
\item[\vn{significant_length}] \Newline
Sets the scale to decide if two length values are significantly different. For example, The
superposition code will not create any super_slave elements that have a length less then this.
%
\item[\vn{sr_wakes_on}] \Newline
Toggle for turning on or off short-range higher order mode wakefield effects.
%
\item[\vn{spin_sokolov_ternov_flipping_on}] \Newline
This determines if the Sokolov-Ternov effect is included in a simulation.  The Sokolov-Ternov
effect\cite{b:barber99} is the self-polarization of charged particle beams due to asymmetric flipping
of a particle's spin when the particle is bent in a magnetic field. Also, spin flipping will {\em
not} be done if spin tracking is off or both radiation damping and excitation are off.
%
\item[\vn{spin_tracking_on}] \Newline
Determines if spin tracking is performed or not.
%
\item[\vn{synch_rad_scale}] \Newline
This parameter is a multiplier for the kick given particles when radiation damping or excitation is
turned on.  This parameter is useful for artificially speeding up (or slowing down) the effect of
radiation.  The default value is one. Values greater than one will give larger kicks and will reduce
the equilibrium settling time.
%
\item[\vn{taylor_order}] \Newline
Cutoff Taylor order of maps produced by \vn{sym_lie_ptc}.
\end{description}

%-----------------------------------------------------------------
\section{PTC_Common_Struct}
\label{s:ptc.common}

The \vn{ptc_common_struct} structure contains a set of global parameters that effect tracking when
PTC is involved. There is only one global instance (\sref{c:param.structs}) of this structure and
this instance has the name \vn{ptc_com}. The components of this structure along with the default
values are:
\begin{example}
  type ptc_common_struct
    max_fringe_order  = 2   ! 2 => Quadrupole.
    complex_ptc_used = True ! Complex PTC code in use? 
    use_totalpath  = False  ! phase space z = time instead of time - ref_time?
    old_integrator = True   ! PTC OLD_INTEGRATOR.
    exact_model    = True   ! PTC EXACT_MODEL.
    exact_misalign = True   ! PTC ALWAYS_EXACTMIS.
    translate_patch_drift_time = True  
  end type
\end{example}

Note: To set the Taylor map order for PTC, set the \vn{taylor_order} parameter of \vn{bmad_com}.

\begin{description}
%
  \index{ptc_exact_model}\index{exact_misalign}
  \item[{parameter[ptc_exact_model]}] \Newline
Deprecated. Replaced by \vn{ptc_com[exact_model]} (\sref{s:bmad.ptc.com}).

The \vn{ptc_exact_model} and \vn{ptc_exact_misalign} switches affect tracking using the \vn{PTC}
library. See \sref{s:integ} for more details.
%
  \index{max_fringe_order}
  \item[{ptc_com[max_fringe_order]}] \Newline
When using PTC tracking (\sref{s:ptc.intro}), the \vn{parameter[ptc_max_fringe_order]} determines
the maximum order of the calculated fringe fields. The default is 2 which means that fringe fields
due to a quadrupolar field. These fields are 3\Rd order in the transverse coordinates.
%
  \index{translate_patch_drift_time}
  \item[{ptc_com[translate_patch_drift_time]}] \Newline
If \vn{translate_patch_drift_time} is set True (the default) the patch in PTC that is setup to
correspond to a \bmad patch is given a reference time offset equal to the \bmad reference time
through the patch. This is generally what is wanted but for a PTC expert who knows what they are
doing and really wants no time offset, \vn{translate_patch_drift_time} can be set False.
  \end{description}

%-----------------------------------------------------------------
\section{Bmad_Com}
\label{s:bmad.ptc.com}
\index{bmad_com}

The parameters of the \vn{bmad_com} instance of the \vn{bmad_common_struct} structure
(\sref{s:bmad.common}) can be set in the lattice file using the syntax
\begin{example}
  bmad_com[parm-name] = value
\end{example}
where \vn{parm-name} is the name of a component of
\vn{bmad_common_struct}. For example:
\begin{example}
  bmad_com[rel_tol_tracking] = 1e-7
\end{example}

A similar situation holds for the \vn{ptc_com} instance of the \vn{ptc_common_struct} structure.

Be aware that setting either a \vn{bmad_com} or \vn{ptc_com} parameter value in a lattice file will
affect all computations of a program even if the program reads in additional lattice files. That is,
setting of \vn{bmad_com} or \vn{ptc_com} components is ``sticky'' and persists even when other
lattice files are read in. There are two exceptions: A program is always free to override settings
of these parameters.  Additionally, a second lattice file can also override the setting made in a
prior lattice file.

%-----------------------------------------------------------------
\section{Space_Charge_Common_Struct}
\label{s:sc.com}
\index{space_charge_common_struct|hyperbf}

The \vn{space_charge_common_struct} structure holds parameters for space charge (including CSR
(\sref{s:csr})) calculations.\footnote
  {
This structure was formally called the \vn{csr_parameter_struct}. The name was changed to reflect
the fact that the structure has parameters for space charge calculations that do not involve CSR.
  }
The setting of the \vn{csr_method} and \vn{space_charge_method} element parameters
(\sref{s:csr.sc.meth}) will also affect space charge calculations as well as the setting of the
\vn{bmad_com} logical \vn{csr_and_space_charge_on} (\sref{s:bmad.common}).

Besides the parameters discussed below, the \vn{csr_and_space_charge_on} parameter of \vn{bmad_com}
(\sref{s:bmad.ptc.com}) must be set True to enable the space charge/CSR calculations. Additionally,
tracking with \vn{CSR} will only be done through elements where the element parameter
\vn{csr_method} (\sref{s:integ}) has been set to something other than \vn{off} and tracking with
space charge will only be done through elements where the element parameter \vn{space_charge_method}
is set to something other than \vn{off}. This is done so that the computationally intensive space
charge and CSR calculations can be restricted to places where the effects are significant.

The space charge / CSR parameter structure has a \vn{type name} of \vn{space_charge_common_struct}
and an \vn{instance name} of \vn{space_charge_com}. This structure has components
\index{space_charge_mesh_size}
\begin{example}
  type space_charge_common_struct 
    ds_track_step = 0                   ! Tracking step size
    dt_track_step = 0                   ! Time based space charge step
    beam_chamber_height = 0             ! Used in shielding calculation.
    cathode_strength_cutoff = 0.01      ! Cutoff for the cathode field calc.
    rel_tol_tracking = 1e-8
    abs_tol_tracking = 1e-10            
    lsc_sigma_cutoff = 0.1              ! Cutoff for the lsc calc. If a bin sigma
                                        !  is < cutoff * sigma_ave then ignore.
    particle_sigma_cutoff = -1          ! Veto particles that are far from the bench with 3D SC.
    n_bin = 0                           ! Number of bins used
    particle_bin_span = 2               ! Longitudinal particle length / dz_bin
    n_shield_images = 0                 ! Chamber wall shielding. 0 = no shielding.
    sc_min_in_bin = 10                  ! Min number of particle needed to compute sigmas.
    space_charge_mesh_size = [32,32,64] ! Mesh size with fft_3d space charge calc.
    csr_3d_mesh_size = [32,32,64]       ! Mesh size for 3D CSR calc.
    print_taylor_warning = T            ! Print Taylor element warning?
    diagnostic_output_file = ""         ! Wake file name
  end type
\end{example}
The values for the various quantities shown above are their default values. 

\begin{description}
\item{ds_track_step} \Newline
The \vn{ds_track_step} parameter sets the nominal longitudinal distance traveled by the bunch
between CSR kicks if the lattice element in which the bunch is being tracked has not set element's
\vn{csr_ds_track} parameter. The actual distance between kicks within a lattice element is adjusted
so that there is an integer number of steps from the element entrance to the element exit. Either
\vn{ds_track_step} or the element's \vn{csr_track_step} must be set to something positive otherwise
an error will result when doing CSR or space charge tracking. Larger values will speed up the
calculation at the expense of accuracy.
%
\item{dt_track_step} \Newline
The \vn{dt_track_step} parameter is used when the \vn{tracking_method} of the lattice element the
bunch is passing through is set to \vn{time_runge_kutta} or \vn{fixed_step_time_runge_kutta}. 
%
\item{beam_chamber_height} \Newline
\vn{beam_chamber_height} is the height of the beam chamber in meters. This parameter is used when
shielding is taken into account.  See also the description of the parameter \vn{n_shield_images}.
%
\item{cathode_strength_cutoff} \Newline
When tracking through an element whose \vn{space_charge_method} is set to \vn{cathode_fft_3d}
(\sref{s:csr.sc.meth}, The value of \vn{cathode_strength_cutoff} is used to determine at how far
from the cathode the cathode image field is included. If the image field is less than
\vn{cathode_strength_cutoff} * bunch field, the image field will be ignored.
%
\item{lsc_sigma_cutoff} \Newline
\vn{lsc_sigma_cutoff} is used in the longitudinal space charge (LSC) calculation and is used to prevent
bins with only a few particles in them to give a large contribution to the kick when the computed
transverse sigmas are abnormally low.
%
\item{n_bin} \Newline
\vn{n_bin} is the number of bins used. The bind width is dynamically adjusted at each kick point so
that the bins will span the bunch length.  This parameter must be set to something positive. Larger
values will slow the calculation while smaller values will lead to inaccuracies and loss of
resolution. \vn{n_bin} should also not be set so large that the average number of particles in a bin
is too small.  ``Typical'' values are in the range 100 --- 1000.
%
\item{particle_bin_span} \Newline
The \vn{particle_bin_span} parameter is the width of a particle's triangular density distribution
(cf.~\sref{s:csr}) in multiples of the bin width. A larger span will give better smoothing of the
computed particle density with an attendant loss in resolution.
%
\item{particle_sigma_cutoff} \Newline
The 3D space charge calculation uses a particle-in-cell algorithm. If there are halo particles far
from the bunch center the grid spacing for the particle-in-cell may become too course. To help remedy
this, particles far from the bunch center may be vetoed by setting \vn{particle_sigma_cutoff} to a 
positive value. When set positive, particles will be ignored in the space charge calc when
\begin{equation}
  \max \left( \frac{|dx|}{\sigma_x}, \frac{|dy|}{\sigma_y}, \frac{|dz|}{\sigma_z} \right)
  > \text{particle_sigma_cutoff}
\end{equation}
where $dx$, $dy$, and $dz$ are the distances along the $x$, $y$, and $z$ axis of the particle from
the bunch centroid, and $\sigma_x$, $\sigma_y$, and $\sigma_z$ are the bunch beam sizes.
%
\item[\vn{\%space_charge_mesh_size}] \Newline
The \vn{%space_charge_mesh_size} sets the size of the grid used when an element's
\vn{space_charge_method} is set to \vn{fft_3d} (\sref{s:csr.sc.meth}). The value of this parameter
is a 3-element array $(n_x, n_y, n_z)$ giving the mesh size in the $x$, $y$, and $z$ directions
respectively. Default values are $(32, 32, 64)$.
%
\item[\vn{\%csr3d_mesh_size}] \Newline
The \vn{%csr3d_mesh_size} sets the size of the grid used when an element's
\vn{csr_method} is set to \vn{steady_state_3d} (\sref{s:csr.sc.meth}). The value of this parameter is a
3-element array $(n_x, n_y, n_z)$ giving the mesh size in the $x$, $y$, and $z$ directions
respectively. Default values are $(32, 32, 64)$.
%
\item{n_shield_images} \Newline
\vn{n_shield_images} is the number of shielding current layers used in the shielding calculation. A
value of zero results in no shielding. See also the description of the parameter
\vn{beam_chamber_height}. The proper setting of this parameter depends upon how strong the shielding
is. Larger values give better accuracy at the expense of computation speed. ``Typical'' values are
in the range 0 --- 5.
%
\item{sc_min_in_bin} \Newline
the \vn{sc_min_in_bin} parameter sets the minimum number of particle in a bin needed to compute the
transverse beam sigmas for that bin. If the number of particles is less than this number, the beam
sigmas are taken to be equal to the beam sigmas of a nearby bin where there are enough particle to
compute the sigma. The beam sigmas are needed for the CS calculation but not need for the CSR
calculation.
%
\item{diagnostic_output_file} \Newline
If set non blank, an output file of this name is created that contains a table of the CSR wake at
each track step (the track step size is set by \vn{ds_track_step}). If tracking is done through
multiple lattice elements, the wake tables for the elements are appended to the file. This file is
useful for visualization of the wake.
\end{description}

Note: \vn{Taylor} map elements (\sref{s:taylor}) that have a finite length cannot be subdivided for
the CSR calculation. \bmad will ignore any \vn{taylor} elements present in the lattice but will
print a warning that it is doing so. To suppress the warning, \vn{print_taylor_warning} should be
set to False.

%-----------------------------------------------------------------
\section{Opti_DE_Param_Struct}
\label{s:de.params}
\index{de optimizer parameters|hyperbf}

\index{opti_de_param_struct}
The Differential Evolution (\vn{DE}) optimizer is used in nonlinear optimization problems. This
optimizer is based upon the work of Storn and Price\cite{b:de}. There are a number of parameters
that can be varied to vary how the optimizer works. These parameters are are contained in a
structure named \vn{opti_de_param_struct}. the instance name is \vn{opti_de_param}.  This structure
has components
\begin{example}
                              Default
  type opti_de_param_struct
    CR             = 0.8    ! Crossover Probability.
    F              = 0.8    !
    l_best         = 0.0    ! Percentage of best solution used.
    binomial_cross = False  ! IE: Default = Exponential.
    use_2nd_diff   = False  ! use F * (x_4 - x_5) term
    randomize_F    = False  !
    minimize_merit = True   ! F => maximize the Merit func.
  end type
\end{example}

The "perturbed vector" is
  v = x_1 + l_best * (x_best - x_1) + F * (x_2 - x_3) + F * (x_4 - x_5)
The last term F * (x_4 - x_5) is only used if \vn{use_2nd_diff} = T.

The crossover can be either "Exponential" or "Binary". 
Exponential crossover is what is described in the paper.
With Exponential crossover the crossover parameters from a contiguous block
and the average number of crossover parameters is approximately
    average crossovers $\sim$ min(D, CR / (1 - CR))
where D is the total number of parameters.
With Binary crossover the probability of crossover of a parameter is 
uncorrelated with the probability of crossover of any other parameter and
the average number of crossovers is
    average crossovers = D * CR

\vn{randomize_F} = True means that the F that is used for a given 
generation  is randomly chosen to be within the range 
[0, 2*F] with average F.

%-----------------------------------------------------------------
\section{Dynamic Aperture Simulations: Aperture_Param_Struct}

\index{dynamic_aperture_struct|hyperbf}
The \vn{dynamic_aperture_struct} is used for dynamic aperture
calculations. This structure has components:
\begin{example}
  type aperture_param_struct
    min_angle = 0
    max_angle = pi
    n_angle   = 9
    n_turn = 100                ! Number of turns a particle must survive
    x_init = 1e-3              ! Initial estimate for horizontal aperture
    y_init = 1e-3              ! Initial estimate for vertical aperture
    accuracy = 1e-5            ! Resolution of bracketed aperture.
  end type
\end{example}

