\chapter{Reading and Writing Lattices}

%----------------------------------------------------------------------------
\section{Reading in Lattices}
\label{s:lat.readin}
\index{lattice files!reading}

There are two subroutines in \bmad to read in a \bmad standard lattice file:
\Hyperref{r:bmad.parser}{bmad_parser} and \Hyperref{r:bmad.parser2}{bmad_parser2}. \vn{bmad_parser}
is used to initialize a \vn{lat_struct} (\sref{c:lat.struct}) structure from scratch using the
information from a lattice file. Unless told otherwise, after reading in the lattice,
\vn{bmad_parser} will compute the 6x6 transfer matrices for each element and this information will
be stored in the \vn{digested file} (\sref{s:digested}) that is created.  Notice that
\vn{bmad_parser} does {\em not} compute any Twiss parameters.

\Hyperref{r:bmad.parser2}{bmad_parser2} is typically used after \vn{bmad_parser} if there is
additional information that needs to be added to the lattice. For example, consider the case where
the aperture limits for the elements is stored in a file that is separate from the main lattice
definition file and it is undesirable to put a \vn{call} statement in one file to reference the
other.  To read in the lattice information along with the aperture limits, there are two
possibilities: One possibility is to create a third file that calls the first two:
\begin{verbatim}
 ! This is a file to be called by bmad_parser
 call, file = ""lattice_file""
 call, file = ""aperture_file""
\end{verbatim}
and then just use \vn{bmad_parser} to parse this third file. The
alternative is to use \vn{bmad_parser2} so that the program code looks
like:
\begin{verbatim}
  ! program code to read in everything
  type (lat_struct) lat
  call bmad_parser ('lattice_file', lat)       ! read in a lattice.
  call bmad_parser2 ('aperture_file', lat)     ! read in the aperture limits.
\end{verbatim}

%----------------------------------------------------------------------------
\section{Digested Files}
\index{digested files}

\index{bmad_parser}
\index{bmad version number}
Since parsing can be slow, once the \vn{bmad_parser} routine has transferred the information from a
lattice file into the \vn{lat_struct} it will make what is called a digested file. A digested file
is an image of the \vn{lat_struct} in binary form. When \vn{bmad_parser} is called, it first looks
in the same directory as the lattice file for a digested file whose name is of the form:
\begin{verbatim}
  ""digested_"" // LAT_FILE 
\end{verbatim}
where \vn{LAT_FILE} is the lattice file name. If \vn{bmad_parser} finds the digested file, it checks
that the file is not out--of--date (that is, whether the lattice file(s) have been modified after
the digested file is made).  \vn{bmad_parser} can do this since the digested file contains the names
and the dates of all the lattice files that were involved. Also stored in the digested file is the
``\bmad \vn{version number}''. The \bmad version number is a global parameter that is increased (not
too frequently) each time a code change involves modifying the structure of the \vn{lat_struct} or
\vn{ele_struct}. If the \bmad version number in the digested file does not agree with the number
current when \vn{bmad_parser} was compiled, or if the digested file is out--of--date, a warning will
be printed, and \vn{bmad_parser} will reparse the lattice and create a new digested file.

\index{taylor map!with digested files}
Since computing Taylor Maps can be very time intensive, \vn{bmad_parser} tries to reuse Taylor Maps
it finds in the digested file even if the digested file is out--of--date. To make sure that
everything is OK, \vn{bmad_parser} will check that the attribute values of an element needing a
Taylor map are the same as the attribute values of a corresponding element in the digested file
before it reuses the map. Element names are not a factor in this decision.

This leads to the following trick: If you want to read in a lattice where there is no corresponding
digested file, and if there is another digested file that has elements with the correct Taylor Maps,
then, to save on the map computation time, simply make a copy of the digested file with the digested
file name corresponding to the first lattice.

\Hyperref{r:read.digested.bmad.file}{read_digested_bmad_file}
\Hyperref{r:write.digested.bmad.file}{write_digested_bmad_file} The digested file is in binary
format and is not human readable but it can provide a convenient mechanism for transporting lattices
between programs. For example, say you have read in a lattice, changed some parameters in the
\vn{lat_struct}, and now you want to do some analysis on this modified \vn{lat_struct} using a
different program.  One possibility is to have the first program create a digested file
\begin{example}
  call write_digested_bmad_file ('digested_file_of_mine', lat)
\end{example}
and then read the digested file in with the second program
\begin{example}
  call read_digested_bmad_file ('digested_file_of_mine', lat)
\end{example}
An alternative to writing a digested file is to write a lattice file
using \vn{write_bmad_lattice_file}

%----------------------------------------------------------------------------
\section{Writing Lattice files}
\label{s:lat.write}
\index{lattice files!MAD files}
\index{MAD}

\index{MAD!MAD-8}
\Hyperref{r:write.bmad.lattice.file}{write_bmad_lattice_file} To create a \bmad lattice file from a
\vn{lat_struct} instance, use the routine
\Hyperref{r:write.bmad.lattice.file}{write_bmad_lattice_file}.  \mad--8, \mad--X, or \vn{SAD}
compatible lattice files can be created from a \vn{lat_struct} variable using the routine
\Hyperref{r:write.lattice.in.foreign.format}{write_lattice_in_foreign_format}:
\begin{example}
  type (lat_struct) lat             ! lattice
  ...
  call bmad_parser (bmad_lat_file, lat)               ! Read in a lattice
  call write_lattice_in_foreign_format ("lat.mad", "MAD-8", lat)  ! create MAD file
\end{example}
Information can be lost when creating a \mad or \vn{SAD} file.  For example, neither \mad nor
\vn{SAD} has the concept of things such as \vn{overlay}s and \vn{group}s.

