\chapter{Taylor Maps}

%-----------------------------------------------------------------
\section{Taylor Maps}
\label{s:taylor.phys}
\index{taylor map|hyperbf}
\index{symplectic integration}

A transport map ${\cal M}: {\cal R}^6 \rightarrow {\cal R}^6$ through an element or a section of a
lattice is a function that maps the starting phase space coordinates $\Bf r(\In)$ to the ending
coordinates $\Bf r(\Out)$
\begin{equation}
  \Bf r(\Out) = {\cal M} \, (\delta \bfr)
\end{equation}
where
\begin{equation}
  \delta \bfr = \bfr(\In) - \bfr_\REF
\end{equation}
$\bfr_\REF$ is the reference orbit at the start of the map around which the map is made. In many
cases the reference orbit is the zero orbit. For a storage ring, the closed orbit is commonly used
for the reference orbit. For a lattice with an open geometry the reference orbit may be the orbit as
computed from some given initial coordinates.

${\cal M}$ in the above equation is made up of six functions ${\cal M}_i: {\cal R}^6 \rightarrow
{\cal R}$. Each of these functions maps to one of the $r(\Out)$ coordinates. Each of these functions
can be expanded in a Taylor series and truncated at some order. Each Taylor series is in the form
\begin{equation}
  r_i(\Out) = \sum_{j = 1}^N \, C_{ij} \, \prod_{k = 1}^6 \, (\delta r_k)^{e_{ijk}}
  \label{rcr}
\end{equation}
Where the $C_{ij}$ are coefficients and the $e_{ijk}$ are integer exponents.  The order of a given
term associated with index $i,j$ is the sum over the exponents
\begin{equation}
  \text{order}_{ij} = \sum_{k = 1}^6 e_{ijk} 
\end{equation}
The order of the entire map is the order at which the map is truncated.

The standard \bmad routine for printing a Taylor map might produce something like this:
\begin{example}
   Taylor Terms:
   Out      Coef             Exponents          Order       Reference
   --------------------------------------------------
    1:     -0.600000000000   0  0  0  0  0  0       0       0.200000000
    1:      1.000000000000   1  0  0  0  0  0       1
    1:      0.145000000000   2  0  0  0  0  0       2
   --------------------------------------------------
    2:     -0.185000000000   0  0  0  0  0  0       0       0.000000000
    2:      1.300000000000   0  1  0  0  0  0       1
    2:      3.800000000000   2  0  0  0  0  1       3
   --------------------------------------------------
    3:      1.000000000000   0  0  1  0  0  0       1       0.100000000
    3:      1.600000000000   0  0  0  1  0  0       1
    3:    -11.138187077310   1  0  1  0  0  0       2
   --------------------------------------------------
    4:      1.000000000000   0  0  0  1  0  0       1       0.000000000
   --------------------------------------------------
    5:      0.000000000000   0  0  0  0  0  0       0       0.000000000
    5:      0.000001480008   0  1  0  0  0  0       1
    5:      1.000000000000   0  0  0  0  1  0       1
    5:      0.000000000003   0  0  0  0  0  1       1
    5:      0.000000000003   2  0  0  0  0  0       2
   --------------------------------------------------
    6:      1.000000000000   0  0  0  0  0  1       1       0.000000000
\end{example}
Each line in the example represents a single \vn{Taylor term}. The Taylor terms are grouped into 6
\vn{Taylor series}. There is one series for each of the output phase space coordinate. The first
column in the example, labeled ``out'', (corresponding to the $i$ index in \Eq{rcr}) indicates the
Taylor series: $1 = x(out)$, $2 = p_x(out)$, etc. The 6 exponent columns give the $e_{ijk}$ of
\Eq{rcr}. In this example, the second Taylor series (\vn{out} = 2), when expressed as a formula,
would read:
\begin{equation}
  p_x(out) = -0.185 + 1.3 \, \delta p_x + 3.8 \, \delta x^2 \, \delta p_z
\end{equation}

\index{taylor map!reference coordinates}
The reference column in the above example shows the input coordinates around which the Taylor map is
calculated. In this case, the reference coordinates where
\begin{equation}
  (x, p_x, y, p_y, z, p_z)_\REF = (0.2, 0, 0.1, 0, 0, 0, 0)
\end{equation}
The choice of the reference point will affect the values of the coefficients of the Taylor map. As
an example, consider the 1-dimension map
\begin{equation}
  x(out) = A \, \sin(k \, \delta x)
\end{equation}
Then a Taylor map to 1\St order is
\begin{equation}
  x(out) = c_0 + c_1 \, \delta x
\end{equation}
where
\begin{align}
  c_1 &= A \, k \, \cos(k \, x_\REF) \\
  c_0 &= A \, \sin(k \, x_\REF)
\end{align}

Taylor maps using complex numbers is also used by \bmad. The output of such maps is similar to the
output for real maps as shown above except that the coefficient has a real and imaginary part.

%-----------------------------------------------------------------
\section{Spin Taylor Map}
\label{s:spin.map}
\index{spin taylor map}

A Taylor map that fully describes spin (\sref{s:spin.dyn}) and orbital motion, would consist of nine
Taylor series (six for the orbital phase space variables and three for the spin components) and each
Taylor series would be a polynomial in nine variables.

To simplify things, \bmad assumes that the effect on the orbital phase space due to the spin
orientation is negligible. That is, Stern-Gerlach effects are ignored. With this assumption, the
orbital part of the map is only dependent on the six orbital variables. This means that
${\pmb\Omega}_{BMT}$ and ${\pmb\Omega}_{EDM}$ in the Thomas-Bargmann-Michel-Telegdi equation
(\sref{tbmt}), are assumed independent of the spin. Thus the spin transport is just a rotation.
\bmad represents this rotation using a quaternion (\sref{s:quat}). Each of the four components of
the quaternion is a Taylor series and the full phase space plus spin transport uses 10 (= 6 orbital
+ 4 spin) Taylor series with each Taylor series only being dependent on the six orbital phase space
coordinates.

Spin transport involves:
\begin{enumerate}[itemsep=-0.6ex, topsep=-0.9ex]
\item Using the six orbital coordinates, evaluate the four spin Taylor series to produce a quaternion $\bfq$.
\item Normalize the quaternion to one: $\bfq \longrightarrow \bfq/|\bfq|$.
\item Rotate the spin vector: $\bfS \longrightarrow \bfq \, \bfS \, \bfq^{-1}$.
\end{enumerate}
The  normalization of the quaternion is needed since the truncation of the map to a finite order will produce errors
in the magnitude of the quaternion.

The standard \bmad routine for printing a spin Taylor map will produce a result that is very similar
as that produced for the orbital phase space. The difference is that there will only be four Taylor
series labeled (\vn{S1}, \vn{Sx}, \vn{Sy}, \vn{Sz}) for the four components of the quaternion. Also the reference orbit
will not be shown (it is exactly the same as the orbital phase space reference orbit).

Note: When tracking a particle's spin through a map, the quaternion used to rotate the spin is
always normalized to one so that the magnitude of the spin will be invariant.

%-----------------------------------------------------------------
\section{Symplectification}
\label{s:symp.method}
\index{symplectification}

If the evolution of a system can be described using a Hamiltonian then
it can be shown that the linear part of any transport map (the Jacobian)
must obey the symplectic condition. If a matrix $\Bf M$ is not symplectic,
Healy\cite{b:healy} has provided an elegant method for finding a symplectic 
matrix that is ``close'' to $\Bf M$. The procedure is as follows:
From $\Bf M$ a matrix $\bfV$ is formed via
\begin{equation}
  \bfV = \Bf S (\Bf I - \Bf M)(\Bf I + \Bf M)^{-1} 
  \label{e:vsimi}
\end{equation}
where $\Bf S$ is the matrix
\begin{equation}
  \Bf S = 
  \begin{pmatrix} 
      0 &  1 &  0 &  0 &  0 &  0 \cr
     -1 &  0 &  0 &  0 &  0 &  0 \cr
      0 &  0 &  0 &  1 &  0 &  0 \cr
      0 &  0 & -1 &  0 &  0 &  0 \cr
      0 &  0 &  0 &  0 &  0 & -1 \cr
      0 &  0 &  0 &  0 & -1 &  0 \cr
  \end{pmatrix}
  \label{s0100}
\end{equation}
$\bfV$ is symmetric if and only if $\Bf M$ is symplectic. In any case,
a symmetric matrix $\Bf W$ near $\bfV$ can be
formed via
\begin{equation}
  \Bf W = \frac{\bfV + \bfV^t}{2}
\end{equation}
A symplectic matrix $\Bf F$ is now obtained by inverting \eq{e:vsimi}
\begin{equation}
  \Bf F = (\Bf I + \Bf S \Bf W) (\Bf I - \Bf S \Bf W)^{-1}
\end{equation}

%-----------------------------------------------------------------
\section{Map Concatenation and Feed-Down}
\label{s:map.concat}

\index{taylor map!feed-down}
Of importance in working with Taylor maps is the concept of
\vn{feed-down}.  This is best explained with an example. To keep the
example simple, the discussion is limited to one phase space
dimension so that the Taylor maps are a single Taylor series. Take the
map $M_1$ from point 0 to point 1 to be
\begin{equation}
  M_1: x_1 = x_0 + 2
  \label{xx2}
\end{equation}
and the map $M_2$ from point 1 to point 2 to be
\begin{equation}
  M_2: x_2 = x_1^2 + 3 \, x_1
  \label{xx3x}
\end{equation}
Then concatenating the maps to form the map $M_3$ from point 0 to point 2
gives
\begin{equation}
  M_3: x_2 = (x_0 + 2)^2 + 3 (x_0 + 2) = x_0^2 + 7 \, x_0 + 10
  \label{xx23x2}
\end{equation}
However if we are evaluating our maps to only 1\St order the map $M_2$
becomes
\begin{equation}
  M_2: x_2 = 3 \, x_1
\end{equation}
and concatenating the maps now gives
\begin{equation}
  M_3: x_2 = 3 (x_0 + 2) = 3 \, x_0 + 6
  \label{x3x23}
\end{equation}
Comparing this to \Eq{xx23x2} shows that by neglecting the 2\Nd order
term in \Eq{xx3x} leads to 0\Th and 1\St order errors in
\Eq{x3x23}. These errors can be traced to the finite 0\Th order term in
\Eq{xx2}. This is the principal of feed--down: Given $M_3$ which is a map
produced from the concatenation of two other maps, $M_1$, and $M_2$
\begin{equation}
  M_3 = M_2(M_1)
\end{equation}
Then if $M_1$ and $M_2$ are correct to n\Th order, $M_3$ will also be
correct to n\Th order as long as $M_1$ has no constant (0\Th order)
term. [Notice that a constant term in $M_2$ does not affect the
argument.]  What happens if we know there are constant terms in our
maps? One possibility is to go to a coordinate system where the
constant terms vanish. In the above example that would mean using the
coordinate $\widetilde x_0$ at point 0 given by
\begin{equation}
  \widetilde x_0 = x_0 + 2
\end{equation}
\index{symplectic integration}

%-----------------------------------------------------------------
\section{Symplectic Integration}
\label{s:symp.integ}

Symplectic integration, as opposed to concatenation, never has
problems with feed--down. The subject of symplectic integration is too
large to be covered in this guide. The reader is referred to the book
``Beam Dynamics: A New Attitude and Framework'' by \'Etienne
Forest\cite{b:forest}. A brief synopsis: Symplectic integration uses
as input 1) The Hamiltonian that defines the equations of motion, and
2) a Taylor map $M_1$ from point 0 to point 1. Symplectic integration
from point 1 to point 2 produces a Taylor map $M_3$ from point 0 to
point 2. Symplectic integration can produce maps to arbitrary
order. In any practical application the order $n$ of the final map is
specified and in the integration procedure all terms of order higher
than $n$ are ignored. If one is just interested in knowing the final
coordinates of a particle at point 2 given the initial coordinates at
point 1 then $M_1$ is just the constant map
\begin{equation}
  M_1: x_1 = c_i
\end{equation}
where $c_i$ is the initial starting point. The order of the integration is set to 0 so that all
non--constant terms are ignored. The final map is also just a constant map
\begin{equation}
  M_3: x_2 = c_f
\end{equation}
If the map from point 1 to point 2 is desired then the map $M_1$ is just set to the identity map
\begin{equation}
  M_1: x_1 = x_0
\end{equation}
In general it is impossible to exactly integrate any non--linear system. In practice, the symplectic
integration is achieved by slicing the interval between point 1 and point 2 into a number of
(generally equally spaced) slices. The integration is performed, slice step by slice step. This is
analogous to integrating a function by evaluating the function at a number of points. Using more
slices gives better results but slows down the calculation. The speed and accuracy of the
calculation is determined by the number of slices and the \vn{order} of the integrator. The concept
of integrator order can best be understood by analogy by considering the trapezoidal rule for
integrating a function of one variable:
\begin{equation}
  \int_{y_a}^{y_b} f(y) \, dy = 
  h \left[ \frac{1}{2} f(y_a) + \frac{1}{2} f(y_b) \right] +
  o(h^3 \, f^{(2)})
\end{equation}
In the formula $h = y_b - y_a$ is the slice width. $0(h^3 \, f^{(2)})$ means that the error of the
trapezoidal rule scales as the second derivative of $f$. Since the error scales as $f^{(2)}$ this is
an example of a second order integrator. To integrate a function between points $y_1$ and $y_N$ we
slice the interval at points $y_2 \ldots y_{N-1}$ and apply the trapezoidal rule to each
interval. The concept of integrator order in symplectic integration is analogous.

The optimum number of slices is determined by the smallest number that gives an acceptable
error. The slice size is given by the \vn{ds_step} attribute of an element (\sref{s:integ}).
Integrators of higher order will generally need a smaller number of slices to achieve a given
accuracy. However, since integrators of higher order take more time per slice step, and since it is
computation time and not number of slices which is important, only a measurement of error and
calculation time as a function of slice number and integrator order will unambiguously give the
optimum integrator order and slice width.  In doing a timing test, it must be remembered that since
the magnitude of any non-nonlinearities will depend upon the starting position, the integration
error will be dependent upon the starting map $M_1$. \bmad has integrators of order 2, 4, and 6
(\sref{s:integ}). Timing tests performed for some wiggler elements (which have strong
nonlinearities) showed that, in this case, the 2\Nd order integrator gave the fastest computation
time for a given accuracy. However, the higher order integrators may give better results for
elements with weaker nonlinearities.
