\begin{thebibliography}{XXXXXXX99}

\bibitem[Abell06]{b:rf.abell}
Dan Abell, ``Numerical computation of high-order transfer maps for rf cavities'',
Phys. Rev. ST Accel. Beams, {\bf 9}, pp. 052001, (2006).

\bibitem[AML]{b:aml}
The Accelerator Markup Language / Universal Accelerator Project web page:
\hfill\break
\hspace*{0.3in}
\url{http://www.lepp.cornell.edu/~dcs/aml/}

\bibitem[Bater64]{b:batterman}
B.~Batterman, and H.~Cole,
``Dynamical Diffraction of X Rays by Perfect Crystals'',
Rev.\ Mod.\ Phys.,{\bf 36}, 3, pp.~ 681--717, (1964).

\bibitem[Berz89]{b:berz}
M. Berz, 
``Differential Algebraic Description of Beam Dynamics to Very High Orders,''
Particle Accelerators, Vol. 24, pp. 109-124, (1989).

\bibitem[Blas94]{b:blasdell}
R.~C.~Blasdell and A.~T.~Macrander, ``Modifications to the 1989 SHADOW
 ray-tracing code for general asymmetric perfect-crystal optics,''
Nuc.\ Instr.\ \& Meth. A {\bf 347}, 320 (1994).

\bibitem[Bmad]{b:bmad.web}
The Bmad web site:
\hfill\break
\hspace*{0.3in} \url{http://www.lepp.cornell.edu/~dcs/bmad}

\bibitem[Rio98]{b:del.rio}
Manuel Sanchez del Rio, ``Ray tracing simulations for crystal optics,''
Proc. SPIE 3448, Crystal and Multilayer Optics, {\bf 230} (1998). 

\bibitem[Brown77]{b:transport.appendix} 
K. L. Brown, F. Rothacker, D. C. Carey, and Ch. Iselin, ``TRANSPORT
Appendix,'' Fermilab, unpublished, (December 1977).

\bibitem[Chao93]{b:chao} 
Alexander Chao, {\em Physics of Collective Beam
Instabilities in High Energy Accelerators}, Wiley, New York (1993). 

\bibitem[Corbett99]{b:corbett}
J. Corbett and Y. Nosochkov, ``Effect of Insertion Devices in SPEAR--3,''
Proc. 1999 Part.\ Acc.\ Conf., p.~238, (1999).

\bibitem[Duff87]{b:leduff}
  J. Le Duff, \emph{Single and Multiple Touschek Effects}.
  Proc. CAS Berlin 1987,
  CERN 89-01,
  1987.

\bibitem[Forest02]{b:ptc}
\'E. Forest, F. Schmidt, E. McIntosh, 
{\it Introduction to the Polymorphic Tracking Code}, 
CERN-SL-2002-044 (AP), and KEK-Report 2002-3 (2002). 
Can be obtained at:
\hfill\break
\hspace*{0.3in}
\url{http://frs.web.cern.ch/frs/report/sl-2002-044.pdf}

\bibitem[Forest06]{b:geo.int}
\'Etienne Forest, `Geometric integration for particle accelerators,''
J. Phys. A: Math. Gen. {\bf 39} (2006) 5321–5377.

\bibitem[Forest88]{b:quad.fringe}
\'E. Forest, J. Milutinovic, 
``Leading Order Hard Edge Fringe Fields Effects Exact in ($1+\delta$) and 
Consistent with Maxwell's Equations for Rectilinear Magnets,''
Nuc. Instrum. and Methods in Phys. Research A {\bf 269}, pp 474-482, (1988).

\bibitem[Forest98]{b:forest}
\'E. Forest, {\em Beam Dynamics: A New Attitude and Framework},
Harwood Academic Publishers, Amsterdam (1998).


\bibitem[Grote96]{b:maduser}
H. Grote, F. C. Iselin, {\it The MAD Program User's Reference Manual},
Version 8.19, CERN/SL/90-13 (AP) (REV. 5) (1996). 
Can be obtained at:
\hfill\break
\hspace*{0.3in}
\url{http://mad.home.cern.ch/mad} 

\bibitem[Healy86]{b:healy}
L. M. Healy, {\it Lie Algebraic Methods for Treating Lattice Parameter
Errors in Particle Accelerators}. Doctoral thesis, University of
Maryland, unpublished, (1986).

\bibitem[Helm73]{b:helm}
R. H. Helm, M. J. Lee, P. L. Morton, and M. Sands, ``Evaluation of Synchrotron
Radiation Integrals,'' IEEE Trans.~Nucl.~Sci. NS-20, 900 (1973).

\bibitem[Hoff06]{b:spin}
G.~Hoffstaetter, {\it Hight-Energy Polarized Proton Beams, A Modern View}, 
Springer. Springer Tracks in Modern Physics Vol~218, (2006).

\bibitem[Iselin94]{b:madphysics}
F. C. Iselin, {\it The MAD program Physical Methods Manual}, 
unpublished, (1994).  Can be obtained at: 
\hfill\break
\hspace*{0.3in}
\url{http://mad.home.cern.ch/mad}

\bibitem[Jowett87]{b:jowett} 
J. M. Jowett, ``Introductory Statistical Mechanics
for Electron Storage Rings,'' AIP Conf. Proc. 153, Physics of Part.\ Acc.,
M. Month and M. Dienes Eds., pp.~864, (1987).

\bibitem[Kohn95]{b:kohn}
V.~G.~Kohn, 
``On the Thcory of Reflectivitlby an X-Ray Multilaler Mirror''
physica status solidi (b), {\bf 187}, 61, (1995).

\bibitem[Press92]{b:nr}
W. Press, B. Flannery, S. Teukolsky, and W. Wetterling, {\em Numerical
Recipes in Fortran, the Art of Scientific Computing}, Second Edition,
Cambridge University Press, New York, (1992). \hfill \break
W. Press, B. Flannery, S. Teukolsky, and W. Wetterling, {\em Numerical
Recipes in Fortran90, the Art of Parallel Scientific Computing}, 
Cambridge University Press, New York, (1996).

\bibitem[Piwin98]{b:piwinski}
Anton Piwinski, \emph{The Touschek Effect in Strong Focusing Storage Rings}.
DESY 98-179, 1998.

\bibitem[Rosen94]{b:rosenzweig}
J. Rosenzweig and L. Serafini, ``Transverse Particle Motion in
Radio--Frequency Linear Accelerators,'' Phys Rev E, Vol. 49, p. 1599,
(1994).

\bibitem[Ruth87]{b:ruth} R. D. Ruth, ``Single-Particle Dynamics in
Circular Accelerators,'' in AIP Conference Proceedings {\bf 153}, {\em
Physics of Particle Accelerators}, pp.~152--235, M. Month and M. Dienes editors,
American Institute of Physics, New York (1987).

\bibitem[SAD]{b:sad} 
D.~Zhou and K.~Oide, ``Maps Used in SAD'' (unpublished).
Also see:
\hfill\break
\hspace*{0.3in} \url{http://acc-physics.kek.jp/SAD/}

\bibitem[Sagan03]{b:wiggler}
D. Sagan, J. Crittenden, and D. Rubin.
``A Symplectic Model for Wigglers,'' Part.\ Acc.\ Conf. (2003).

\bibitem[Sagan99]{b:coupling}
D. Sagan and D. Rubin ``Linear Analysis of Coupled Lattices,''
Phys.\ Rev.\ ST Accel.\ Beams {\bf 2}, 074001 (1999).
\hfill\break
\hspace*{20pt} 
\url{http://link.aps.org/doi/10.1103/PhysRevSTAB.2.074001}

\bibitem[Sagan06]{b:csr}
D. Sagan, ``An Efficient Formalism for Simulating the Longitudinal Kick from Coherent 
Synchrotron Radiation,'' Proc. Europ.\ Part.\ Accel.\ Conf. p. 2829 --- 31 (2006).

\bibitem[Silenko08]{b:silenko}
Alexander Silenko, ``Equation of Spin Motion in Storage Rings in a Cylindrical Coordinate System,''
Phys.\ Rev.\ ST Accel.\ Beams, {\bf 9} pp. 034003, (2006).


\bibitem[Storn96]{b:de}
R.~Storn, and K.~V.~Price, ``Minimizing the real function of the
ICEC'96 contest by differential evolution'' IEEE conf. on Evolutionary
Computation, 842-844 (1996).

\bibitem[Stoltz02]{b:boris}
P. H. Stoltz and J. R. Cary, ``Efficiency of a Boris--like Integration
Scheme with Spatial Stepping,'' Phys.\ Rev.\ Special Topics ---
Accel. \& Beams {\bf 5}, 094001 (2002).

\bibitem[Talman87]{b:talman} R. Talman, ``Multiparticle Phenomena and
Landau Damping,'' in AIP Conf.\ Proc.  {\bf 153}, {\em Physics of
Particle Accelerators}, pp.~789--834, M. Month and M. Dienes editors,
American Institute of Physics, New York (1987).

\bibitem[Tao]{b:tao}
D. Sagan, J. Smith, {\it The Tao Manual}.
Can be obtained at: \hfill\break
\hspace*{0.3in}
\url{http://www.lepp.cornell.edu/~dcs/bmad/tao_entry_point.html}

\bibitem[Venturini98]{b:vent.map} 
M Venturini, D Abell, A Dragt, 
``Map Computation from Magnetic Field Data and Application to the LHC
High-Gradient Quadrupoles,'' 
Proc.\ 1998 International Comp.\ Accel.\ Phys.\ Conf.\ (ICAP '98),
{\bf 980914}, 184 (1998).

\bibitem[Rauben91]{b:tol}
T. Raubenheimer,
``Tolerances to Limit the Vertical Emittance in Future Storage Rings'', 
Particle Accelerators, 1991, {\bf 36}, pp.75-119. 
SLAC-PUB-4937 Rev., (1991).

\bibitem[Wiede99]{b:wiedemann}
H. Wiedemann, {\em Particle Accelerator Physics}, Springer, New York, 3rd Edition (2007). 

\bibitem[Wolski06]{b:wolski.coupling}
A.~Wolski,  ``Alternative approach to general coupled linear optics,''
Phys. Rev. ST Accel. Beams 9, 024001 (2006).

\bibitem[Wyckoff65]{b:wyckoff}
R. W. G. Wyckoff, {\em Crystal Structures}, Interscience Publ. (1965).

\bibitem[Schoon11]{b:xraylib} 
T. Schoonjans et al. ``The xraylib library for X-ray-matter
interactions. Recent developments,'' Spectrochimica Acta Part B: Atomic
Spectroscopy {\bf 66}, pp. 776-784 (2011).

\index{XSIF!reference}
\bibitem[Tenen01]{b:xsif}
P. Tenenbaum, ``LIBXSIF, A Stand alone Library for Parsing the Standard 
Input Format,'' Proc.\ 2001 Part.\ Acc.\ Conf.\ p. 3093 --- 95 (2001).
Documentation at
\hfill\break
\hspace*{0.3in} \url{http://www-project.slac.stanford.edu/lc/ilc/TechNotes/LCCNotes/PDF/LCC-0060%20rev.1.pdf}

\end{thebibliography}
