\chapter{Introduction}
\label{c:introduction}

%------------------------------------------------------------------------
\section{Overview}
\label{s:overview}

\tao stands for ``Tool for Accelerator Optics''. \tao is a general purpose program for simulating
high energy particle beams in accelerators and storage rings. The simulation engine that \tao uses
is the \bmad software library\cite{b:bmad}. \bmad was developed as an object-oriented library so
that common tasks, such as reading in a lattice file and particle tracking, did not have to be coded
from scratch every time someone wanted to develop a program to calculate this, that or whatever.

After the development of \bmad, it became apparent that many simulation programs had common needs:
For example, plotting data, viewing machine parameters, etc. Because of this commonality, the \tao
program was developed to reduce the time needed to develop a working programs without sacrificing
flexibility. That is, while the ``vanilla'' version of the \tao program is quite a powerful
simulation tool, \tao has been designed to be easily customizable so that extending \tao to solve
new and different problems is relatively straight forward.

More information, including the most up--to--date version of this manual, can be found at the \bmad
web site at:
\begin{example}
  classe.cornell.edu/bmad
\end{example}

Errors and omissions are a fact of life for any reference work and comments from you, dear reader,
are therefore most welcome. Please send any missives (or chocolates, or any other kind of
sustenance) to:
\begin{example}
  David Sagan <dcs16@cornell.edu>
\end{example}

It is my pleasure to express appreciation to people who have contributed to this effort. To Scott
Berg, Michael Ehrlichman, Chris Mayes, and Jeff Smith for bug reports, suggestions, code
improvements, Etc. To John Mastroberti and Kevin Kowalski for their work on a graphics user
interface and associated plotting, And last but not least thanks also must go to Dave Rubin and
Georg Hoffstaetter for their help, support, and patience.

%------------------------------------------------------------------------
\section{Manual Organization}

This manual is divided into two parts. Part I is the reference section which defines the terms used
by \tao, discusses \tao commands, etc. Part II is a programmer's guide which shows how to extend
\tao's capabilities including interfacing to Python and how to incorporate custom calculations into
\tao.

%------------------------------------------------------------------------
\section{Bmad and Tao Tutorial}

\vspace{0.1in}
The \tao manual is organized as reference guide and so does not do a good job of instructing the
beginner as to how to use \tao. For that there is an introduction and tutorial on \bmad and \tao
concepts that can be downloaded from the \bmad web page. Go to either the \bmad or \tao manual pages
and there will be a link for the tutorial.

%------------------------------------------------------------------------
\section{Other Bmad based programs}
\label{s:other}

\tao is not the only program based upon \bmad. There are other programs which are specialized for
certain computations such as long-term tracking, beam break-up instability, multi-objective
optimization, etc. The discussion of these programs is beyound the scope of this manual. More
information along with instructions for obtaining the programs can be obtained from the \bmad web site (\sref{s:overview}).

