\chapter{MAD/SAD/PTC Lattice Conversion}
\label{c:lat.convert}
\index{conversion to other lattice formats}

%-----------------------------------------------------------------------------
\section{MAD Conversion}
\label{s:mad.convert}
\index{MAD!conversion}

%-----------------------------------------------------------------------------
\subsection{Convert MAD to Bmad}
\label{s:mad.bmad.uap}

Python scripts to convert from MAD8 and MADX are available at:
\begin{example}
  util_programs/mad_to_bmad
\end{example}
Due to differences in language definitions, conversions must be done with some care. The following
differences should be noted:
  \begin{itemize}
  \item
\bmad, unlike \mad, does not have any ``action'' commands. An action command is a command that makes
a calculation. Examples include \mad's \vn{SURVEY} and \vn{TWISS} commands.
  \item
In \bmad all variables must be defined before being used (\sref{s:arith}) while \mad will simply take
a variable's value to be zero if it is not defined.
  \item
\bmad, unlike \mad, does not allow variable values to be redefined.
  \end{itemize}

%-----------------------------------------------------------------------------
\subsection{Convert Bmad to MAD}
\label{s:bmad.mad}

\index{wiggler!conversion to MAD}
\index{sol_quad!conversion to MAD}
To convert to MAD8 or MADX, there is a \bmad conversion routine called
\Hyperref{r:write.lattice.in.foreign.format}{write_lattice_in_foreign_format}.  Since \mad does not
have a \vn{wiggler} or a \vn{sol_quad} element, this conversion routine makes ``equivalent''
substitution. For a \vn{sol_quad}, the equivalent substitution will be a drift-matrix-drift series
of elements. For a \vn{wiggler}, a series of bend and drift elements will be used (the program can
also use a drift-matrix-drift model here but that is not as accurate). The bends and drifts for the
\vn{wiggler} model are constructed so that the global geometry of the lattice does not
change. Additionally the bends and drifts are constructed to most nearly match the wiggler's
\begin{example}
  Transfer matrix
  $I_2$ and $I_3$ synchrotron radiation integrals (\sref{s:synch.ints})
\end{example}
Note that the resulting model will not have the vertical cubic
nonlinearity that the actual wiggler has.

The \vn{write_lattice_in_foreign_format} routine is embeded in the program
\vn{util_programs/bmad_to_mad_or_sad}.

%-----------------------------------------------------------------------------
%\subsection{Convert MAD-X to Bmad Via PTC}
%\label{s:mad.bmad.ptc}
%
%\mad-X lattices can be converted to \bmad by first using \mad-X to create a PTC ``flat''
%file and then using \bmad to convert the flat file to a \bmad lattice file. 
%
%The advantage of converting via PTC rather than using UAP (\sref{s:mad.bmad.uap}) is that
%one does not have to worry about removing action commands from the \mad-X file. The
%disadvantages are:
%  \begin{description}
%  \item 
%the conversion does not preserve any of the look and feel of the original \mad-X file. 
%  \item
%the conversion via PTC does not work with \mad8 lattices.
%  \item
%\mad-X does not correctly handle \vn{elseparator} and \vn{matrix} elements. These
%elements become drifts in a flat file.
%  \item
%\mad-X does not transfer beginning Twiss and orbit information to the flat file needed for open lattices.
%  \item
%\mad-X does not transfer the zeroth (dipole) order multipoles of a multipole element to the flat file.
%  \end{description}
%
%First, take a \mad-X lattice file and append the following to it:
%\begin{example}
%  use, period = LINE_NAME_HERE;
%  ptc_create_universe;
%  ptc_create_layout, model = 1, method = 2, nst = 1; 
%  !! ptc_create_layout, model=1, method=2, nst=1, exact;  ! Use this if exact is wanted
%  ptc_script, file="create_flat.ptc";
%\end{example}
%Substitute the actual sequence or line name for ``LINE_NAME_HERE''.
%
%Second, create a file called \vn{create_flat.ptc} with the following in it:
%\begin{example}
%  select layout
%  1
%  print flat file
%  XXX.flat
%  return
%\end{example}
%The flat file name \vn{XXX.flat} can be changed if desired.
%
%Now run \mad-X with the lattice file. A flat file will be created. This flat file can
%converted to \bmad as documented in \sref{s:ptc.convert}

%---------------------------------------------------------------------------
%\section{PTC Conversion}
%\label{s:ptc.convert}
%\index{ptc!conversion}
%
%For a programmer, 
%For conversion from \bmad to PTC there are two possibilites

%---------------------------------------------------------------------------
\section{SAD Conversion}
\label{s:sad.convert}
\index{SAD}

Conversion from \vn{SAD}\cite{b:sad} to \bmad is accomplished using the Python script
\begin{example}
  util_programs/sad_to_bmad/sad_to_bmad.py
\end{example}
Currently, a converter from \bmad to SAD is planned but has not been implemented.

Currently, the following restrictions on SAD lattices apply:
  \begin{itemize}
  \item
SAD must elements cannot have an associated RF field
  \item
Misalignments in a \vn{sol} element with \vn{geo} = 1 cannot be handled.
  \end{itemize}

%----------------------------------------------------------------------------
\section{Translation Using the Universal Accelerator Parser}
\label{s:aml}
\index{Accelerator Markup Language (AML)}
\index{Universal Accelerator Parser (UAP)}

\index{MAD}
The \vn{Accelerator Markup Language} (\vn{AML}) / \vn{Universal Accelerator Parser} (\vn{UAP})
project\cite{b:aml} is a collaborative effort with the aim of 1) creating a lattice format (the
\vn{AML} part) that can be used to fully describe accelerators and storage rings, and 2) producing
software (the \vn{UAP} part) that can parse \vn{AML} lattice files. A side benefit of this project
is that the \vn{UAP} code has been extended to be able to translate between \vn{AML}, \bmad, \mad-8,
and \mad-X

The program \vn{translate_driver} which comes with the \vn{UAP} code can be used for conversions. To
get help with how to run this program use the command
\begin{example}
  path-to-uap-dir/bin/translate_driver -help
\end{example}
Example:
\begin{example}
  path-to-uap-dir/bin/translate_driver -constants_first -bmad xxx.madx
\end{example}
This will convert a \mad-X file \vn{xxx.madx} to \bmad format.
