\begin{thebibliography}{99}

\bibitem{b:coupling}
D. Sagan and D. Rubin ``Linear Analysis of Coupled Lattices,''
Phys.\ Rev.\ ST Accel.\ Beams {\bf 2}, 074001 (1999).

\bibitem{b:maduser}
H. Grote, F. C. Iselin, {\it The MAD Program User's Reference Manual},
Version 8.19, CERN/SL/90-13 (AP) (REV. 5) (1996). Can be obtained at:
$<$http://mad.home.cern.ch/mad$>$ (under MAD version 8).

\bibitem{b:madphysics}
F. C. Iselin, {\it The MAD program Physical Methods Manual}, 
unpublished, (1994).  Can be obtained at: $<$http://mad.home.cern.ch/mad$>$
(under MAD version 8).

\bibitem{b:scr}
D. Sagan, J. Crittenden, and D. Rubin.
``A Symplectic Model for Wigglers," Part.\ Acc.\ Conf. (2003).

\bibitem{b:healy}
L. M. Healy, {\it Lie Algebraic Methods for Treating Lattice Parameter
Errors in Particle Accelerators}. Doctoral thesis, University of
Maryland, unpublished, (1986).

\bibitem{b:rosenzweig}
J. Rosenzweig and L. Serafini, ``Transverse Particle Motion in
Radio--Frequency Linear Accelerators,'' Phys Rev E, Vol. 49, p. 1599,
(1994).

\bibitem{b:nr}
W. Press, B. Flannery, S. Teukolsky, and W. Wetterling, {\em Numerical
Recipes in Fortran, the Art of Scientific Computing}, Second Edition,
Cambridge University Press, New York, (1992).

\bibitem{b:nr.f90}
W. Press, B. Flannery, S. Teukolsky, and W. Wetterling, {\em Numerical
Recipes in Fortran90, the Art of Parallel Scientific Computing}, 
Cambridge University Press, New York, (1996).


\bibitem{b:boris}
P. H. Stoltz and J. R. Cary, ``Efficiency of a Boris--like Integration
Scheme with Spatial Stepping,'' Phys.\ Rev.\ Special Topics ---
Accel. \& Beams {\bf 5}, 094001 (2002).

\bibitem{wiedemann}
H. Wiedemann, {\em Particle Accelerator Physics}, Springer, New York, (1999). 

\bibitem{b:forest}
E. Forest, {\em Beam Dynamics: A New Attitude and Framework},
Harwood Academic Publishers, Amsterdam (1998).

\end{thebibliography}
