\chapter{The \bmad\ Distribution}

\section{Libraries in the \bmad\ Distribution}

When installing \bmad\ on a computer what one gets is not only the \bmad\
subroutine library but additionally subsidiary libraries upon which
subroutines in the \bmad\ library depend. There are 5 other libraries
that are used: cesr\_utils, forest, numerical\_recipes, pgplot, and dcslib.
\begin{description}
\item[cesr\_utils] This is a small low level library that primarily defines 
the precision that \bmad\ works at (see below) and defines the physical
and mathematical constants (pi, c\_light, etc.) that \bmad\ knows
about.
\item[dcslib] This library defines a set of miscellaneous helper routines. 
Routines include spline fitting, Gaussian random number generation,
etc. The library name comes from its creator.
\item[forest] This is the FPP/PTC 
(Fully Polymorphic Package/ Polymorphic Tracking Code) library of
Etienne Forest that handles Taylor maps to any arbitrary order (this
is also known as Truncated Power Series Algebra (TPSA)). The FPP part
handles the TPSA and the PTC part does the physics of tracking through
elements using Lie Algebra with a Hamiltonian.  \bmad\ uses this
software to crate Taylor Maps, track particles, etc.  FPP/PTC is a
very general package and \bmad\ only makes use of a small part of its
features. For more information see the FPP/PTC web site at
\begin{verbatim} 
    <http://bc1.lbl.gov/CBP_pages/educational/TPSA_DA/Introduction.html>
\end{verbatim}
\item[recipes] Numerical Recipes is a set of subroutines for doing 
scientific computing including Runge--Kutta integration, FFT's,
interpolation and extrapolation, etc., etc. The writeup for this
library is the book ``Numerical Recipes, The Art of Scientific
Computing''\cite{?}. For \bmad\ this library has been modified to handle
both single and double precision reals.
\item[pgplot] The PGPLOT Graphics Subroutine Library is a Fortran or 
C-callable, device-independent graphics package for making simple
scientific graphs.  One
disadvantage of PGPLOT is that it is not the most friendly software
for the programmer. To remidy this, there is a set of Fortran90
wrapper subroutines called quick\_plot. The quick\_plot suite is part
of the dcslib library. More information may be obtained from the PGPLOT
web site at 
\begin{verbatim}
    <http://www.astro.caltech.edu/~tjp/pgplot>.
\end{verbatim}

\end{description}

\section{Precision}

\bmad\ comes in two flavors: One where the real numbers are single
precision and a version with double precision reals. Which version you
are working with is controlled by the parameter \rp\ (Real Precision)
which is defined in cesr\_utils. [Note: For compatibility with older
programs the parameter \rdef\ is defined to be equal to \rp.]  On most
machines single precision has \rp\ = 4 and double precision has \rp\ =
8. Normally the double precision version is used since round-off
errors can be significant in some calculations. Long--term tracking is
an example where the single precision version is not adequate. 

To define your variables with the correct precision use the syntax
{\it real(rp)}. For example:
\begin{verbatim}
    real(rp) var1, var2, var3
\end{verbatim}
When you want to define a literal constant, for example to pass an
argument to a subroutine, add the suffix {\it \_rp} to the end of the
constant. For example: {\it 2.0\_rp} is equivalent to {\it 2.0D0} if
\rp is defined to be double precision. Notice that this is not
equivalent to {\it 2\_rp} which defines an integer (not a real) constant.


