\chapter{Lattice File Overview}
\index{Lattice files|textbf}

%---------------------------------------------------------------------------
\section{Lattice Files}
\label{s:lattice_file_formats}

\index{XSIF}
\index{Attribute|textbf}
A lattice file (or files) defines the sequence of elements that a
particle will travel through along with the attributes of the elements
(their lengths, their strengths, etc.). 
The \bmad software library comes with routines to read in (parse) a
lattice definition files. There are two input formats that
\bmad understands: The \bmad standard format and
XSIF\cite{b:xsif}. XSIF stands for ``Extended Standard Input Format.''
XSIF is essentially a subset of the MAD input format. XSIF
Documentation
can be found at:
\begin{example}
  http://www-project.slac.stanford.edu/
              lc/ilc/TechNotes/LCCNotes/PDF/LCC-0060%20rev.1.pdf
\end{example}
\index{MAD}
One point that is not covered in the XSIF documentation is that for a
\vn{MATRIX} element, unlike \mad, the \vn{R$ii$} terms (the diagonal
terms of the linear matrix) are not unity by default. Thus
\begin{example}
  m: matrix
\end{example}
will give a matrix with all elements being zero.

Throughout this manual, unless explicitly
stated otherwise, the lattice format under discussion
will always pertain to the \bmad standard format.

\index{MAD!lattice format}
The syntax that a \bmad standard lattice file must conform to is
modeled after the lattice input format of the \mad program.
Essentially, a \bmad lattice file is similar to a \mad lattice file
except that a \bmad file has no ``action'' commands (action commands
tell the program to calculate the Twiss parameters, do tracking,
etc.).  Interacting with the user to determine what actions a program
should take is left to the program and is not part of \bmad (although
\bmad, of course, provides the routines to perform many
calculations). A program is not required to use the \bmad parser
routine but if it does the following chapters describe how to
construct a valid lattice file.

%---------------------------------------------------------------------------
\section{File Example and Syntax}
\index{Lattice files!statement syntax}

The following (rather silly) example shows how an input file can look like
\begin{example}
  ! This is a comment
  parameter[beam_energy] = 5e9                   ! Parameter definition
  pa1 = sin(3.47 * pi / c_light)                 ! Constant definition
  bend1: sbend, type = "arc bend", l = 2.3, &    ! An element definition
      g = 2*pa1, tracking_method = bmad_standard
  bend2: bend1, l = 3.4                          ! Another element def
  bend2[g] = 1 / 37.5                            ! Redefining an attribute
  ln1: line = (ele1, ele2, ele3)                 ! A line definition
  ln2: line = (ln1, ele4, ele5)                  ! Lines can contain lines
  arg_ln(a, b) = (ele1, a, ele2, b)              ! A line with arguments.
  use, ln2                                       ! Which line to use for the lattice
\end{example}

A \bmad standard lattice file is free format. A \bmad lattice file consists
of a sequence of statements. Normally a statement occupies a single
line in the file. Several statements may be placed on the same line by
inserting a semicolon (;) between them. A long statement can occupy
multiple lines by putting an ampersand (\&) at the end of each line of
a statement except for the last line of the statement. An
exclamation mark (!) denotes a comment and the exclamation mark and
everything after the exclamation mark on the line are ignored. 
\bmad is case insensitive. All names are converted to uppercase except
strings (\sref{s:var_types}).

%----------------------------------------------------------------------------
\section{Digested Files}
\label{s:lattice_files}
\index{Lattice files!digested files}

Normally the \bmad parser routine will create what is called a
``digested file'' after it has parsed a lattice file so that when a
program is run and the same lattice file is to be read in again, to save
time, the digested file can be used to load in the lattice information.
This digested file is in binary format and is not human readable. The
digested file will contain the transfer maps for all the elements. 
Using a digested file can save considerable time if some of the
elements in the lattice need to have Taylor maps computed.
(this occurs typically with map--type wigglers).

\bmad creates the digested file in the same area as the lattice file.
If \bmad is not able to create a digested file (typically because it
does not have write permission in the directory), an error message will
be generated but otherwise program operation will be normal.

Digested files can also be used for easy transport of lattices between
programs or between sessions of a program. For example, using one
program you might read in a lattice, make some adjustments (say to model
shifts in magnet positions) and then write out a digested version of the
lattice. This adjusted lattice can now be read in by another program to
do some custom analysis.

%---------------------------------------------------------------------------
\section{Element Sequence Definition}

\index{Line}
\index{List}
The creation of the element sequence that defines a lattice involves
constructs called \vn{lines} and \vn{lists} which define subsequencies
that can be concatenated together. For example:
\begin{example}
  l3: line = (l1, l2)   ! Concatenate two lines
  l1: line = (a, b, c)  ! Line with 3 elements
  l2: line = (a, z)     ! Another line 
  use, l3               ! Use l3 as the lattice definition.
  expand_lattice        ! Expand all lines and form the lattice.
\end{example}
In this case the lattice would be
\begin{example}
  (a, b, c, a, z)
\end{example}
\vn{Lines} and \vn{lists} can be defined in any order. Thus in the
above example \vn{l3} was defined in terms of \vn{l1} and \vn{l2}
before these lines where defined. See \cref{c:sequence} for more details.

\index{Superimpose}
The \vn{superimpose} construct allows elements to be placed in a
lattice at a definite longitudinal position. What happens is that
after a lattice is expanded, there is a reshuffling of the elements to
accommodate any new superimpose elements. See \sref{s:super} for more
details.

%---------------------------------------------------------------------------
\section{Lattice Elements}

\index{MAD!lattice format}
\index{Attribute}
The syntax for defining a lattice element roughly follows \mad:
\begin{example}
  label: keyword [, attributes]
\end{example}
\vn{Overlay} and \vn{Group} elements have a slightly different syntax
\begin{example}
  label: keyword = \{ list \}, master-attribute [= value] [, attributes]
\end{example}
and \vn{I_Beam} elements have the syntax
\begin{example}
  label: keyword = \{ list \} [, attributes]
\end{example}  
For example:
\begin{example}
  q01w: quadrupole, type = "A String", l = 0.6, tilt = pi/2
  h10e: overlay = \{ b08e, b10e \}, hkick
\end{example}

\cref{c:elements} gives a list of elements with their attributes.

%---------------------------------------------------------------------------
\section{Element Attributes}
\index{Element attribute|textbf}

Any lattice element has various attributes like its name, its length,
its strength, etc. The values of element attributes can be specified
when the element is defined. For example:
\begin{example}
  b01w: sbend, l = 6.0, rho = 89.0 ! Define an element with attributes.
\end{example}
After an element's definition, an individual attribute may be referred
to using the syntax
\begin{example}
  element-name[attribute-name]
\end{example}
Element attributes can be set or used in an algebraic expression:
\begin{example}
  bo1w[roll] = 6.5                  ! Set an attribute value.
  b01w[l] = 6.5                     ! Change an attribute value.
  b01w[l] = b01w[rho] / 12          ! OK to reset an attribute value.
  my_const = b01[rho] / b01[l]      ! Use of attribute values in an expression.
\end{example}
Notice that there can be no space between the element name and the
\vn{[} opening bracket.  When setting an attribute value if more than
one element has the \vn{element-name} then {\it all} such elements
will be set.  When setting an attribute value if \vn{element-name} is
the name of a type of element then all elements of that type will be
set. For example
\begin{example}
  rfcavity[tracking_method] = symp_lie_ptc ! Set all RFcavity elements.
\end{example}

\cref{c:elements} gives a list of elements with their attributes.

%---------------------------------------------------------------------------
\section{Variable Types}
\label{s:var_types}
\index{Arithmetic expressions!variables}

\index{Logicals|textbf}
There are five types of variables in \bmad: Reals, Integers, Switches,
Logicals (booleans), and Strings. Acceptable logical values are
\begin{example}
  .true.  .false.
   true    false
   t       f
\end{example}
For example
\begin{example}
  rf1[is_on] = False
\end{example}

\index{Strings|textbf}
String literals can be quoted using double quotes (") or single quotes ('). 
If there are no
blanks or commas within a string, the quotes can be omitted. For example:
\begin{example}
  Q00W: Quad, type = "My Type", alias = Who_knows, &
                                  descrip = "Only the shadow knows"
\end{example}
Unlike most everything else, strings are not converted to uppercase.

\index{Switches|textbf}
Switches are variables that take discrete values. For example:
\begin{example}
  beam, particle = positron          
  q01w: quad, tracking\_method = bmad\_standard 
\end{example}
The name ``switch'' can refer to the variable (for example,
\vn{tracking_method}) or to a value that it can take (for example,
\vn{bmad_standard}). The name ``method'' is used interchangeably with switch.

%---------------------------------------------------------------------------
\section{Arithmetic Expressions}
\index{Arithmetic expressions} 

Arithmetic expressions can be used in a place where a real value is required.
The standard operators are defined: \hfil\break
\hspace*{0.15in}
\begin{tabular}{ll}
  $a + b$           & Addition        \\
  $a - b$           & Subtraction     \\
  $a \, \ast \, b$  & Multiplication  \\
  $a \; / \; b$     & Division        \\
  $a \, \land \, b$ & Exponentiation  \\
\end{tabular}
\bmad also has a set of intrinsic functions. A list of these is given
in \sref{s:functions}.

\index{Arithmetic expressions!constants}
Literal constants can be entered with or without a decimal point. An
exponent is marked with the letter E. For example
\begin{example}
  1, 10.35, 5E3, 314.159E-2
\end{example}
Symbolic constants can be defined using the syntax
\begin{example}
  parameter_name = expression
\end{example}
\index{MAD!syntax compatibility with BMAD}
Alternatively, to be compatible with \mad, using ``:='' instead of ``='' is accepted
\begin{example}
  parameter_name := expression
\end{example}
Examples:
\begin{example}
  my_const = sqrt(10.3) * pi^3
  abc     := my_const * 23
\end{example}
\index{MAD!delayed substitution}
Unlike \mad, \bmad uses immediate substitution so that all constants
in an expression must have been previously defined. For example, the
following is not valid:
\begin{example}
  abc      = my_const * 23      ! No: my_const needs to be defined first.
  my_const = sqrt(10.3) * pi^3
\end{example}
here the value of \vn{my_const} is not known when the line ``\vn{abc}
= $\ldots$'' is parsed. Once
defined, symbolic constants cannot be redefined. For example:
\begin{example}
  my_const = 1
  my_const = 2  ! No: my_const cannot be redefined.
\end{example}

%---------------------------------------------------------------------------
\section{Intrinsic functions}
\label{s:functions}
\index{Intrinsic functions}

The following intrinsic functions are recognized by \bmad: \hfil\break
\hspace*{0.15in}
\begin{tabular}{ll}
  \vnif{sqrt}(x)      & Square Root    \\
  \vnif{log}(x)       & Logarithm      \\
  \vnif{exp}(x)       & Exponential    \\
  \vnif{sin}(x)       & Sine           \\
  \vnif{cos}(x)       & Cosine         \\
  \vnif{tan}(x)       & Tangent        \\
  \vnif{asin}(x)      & Arc sine       \\
  \vnif{acos}(x)      & Arc cosine     \\
  \vnif{atan}(x)      & Arc Tangent    \\
  \vnif{abs}(x)       & Absolute Value \\
  \vnif{ran}()        & Random number between 0 and 1 \\
  \vnif{ran_gauss}()  & Gaussian distributed random number \\
\end{tabular}

\index{Ran_seed!parameter statement}
\vn{ran_gauss} is a Gaussian distributed random number with unit RMS. 
Both \vn{ran} and \vn{ran_gauss} use a seeded random number generator. 
To choose the seed set 
\begin{example}
  parameter[ran_seed] = Integer
\end{example}
A \vn{value} of zero will set the seed using the system clock so that
different sequences of random numbers will be generated each time a
program is run.  The default behavior if \vn{parameter[ran_seed]} is
not present is to use the system clock for the seed.

\index{Expand_lattice!statement}
If an element is used multiple times in a lattice, and if \vn{ran} or
\vn{gauss_ran} is used to set an attribute value of this element, then
to have all instances of the element have different attribute values
the setting of the attribute must be after the lattice has been
expanded. For example:
\begin{example}
  a: quad 
  a[x_offset] = 0.001*ran_gauss()
  my_line: line = (a, a)
  use, my_line
\end{example}
Here, because \bmad does immediate evaluation, the \vn{x_offset}
values for \vn{a} gets set in line 2 and so both copies of \vn{a} in
the lattice get the same value. This is probably not what is wanted.
On the other hand if the attribute is set after lattice expansion:
\begin{example}
  a: quad 
  my_line: line = (a, a)
  use, my_line
  expand_line
  a[x_offset] = 0.001*ran_gauss()
\end{example}
Here the two \vn{a} elements in the lattice get different values for
\vn{x_offset}.
