\SECTION{Overview}

\bmad\ (Otherwise known as ``Baby MAD" or ``Better MAD" or just plain
``Be MAD!")  is an subroutine library used for relativistic
charged--particle dynamics simulations in high energy accelerators and
storage rings. \bmad\ has been developed at Cornell University's
Laboratory for Elementary Particle Physics and has been in use since
the 1990's. 

Prior to \bmad, simulation programs were written almost from scratch to perform
calculations that were beyond the capability of existing, generally
available software. This practice was inefficient, leading to much
duplication of effort.  Since it was time consuming to write
simulations, needed calculations where not being done.  As a response,
the \bmad\ subroutines, using an object oriented approach and written
in Fortran90, were developed to:
\begin{Itemize}
\item Cut down on the time needed to develop programs.
\item Minimize computation times.
\item Cut down on programming errors, 
\item Provide a simple mechanism for lattice function calculations
from within programs (control system or otherwise).
\item Provide a flexible yet powerful lattice input format.
\item Standardize sharing of lattice information between 
programs.
\end{Itemize}

\bmad has a wide range of routines to do many things.  \bmad\ can be
used to study both single and multi--particle beam dynamics.  It has
routines to track both particles and macroparticles. \bmad has various
tracking algorithms including Runge--Kutta and symplectic (Lie
algebraic) integration.  Wakefields, and radiation excitation and
damping can be simulated. \bmad\ has routines for calculating transfer
matrices, emittances, Twiss parameters, dispersion, coupling, etc. The
elements that \bmad\ knows about include quadrupoles, RF cavities
(both storage ring and LINAC accelerating types), solenoids, dipole
bends, etc. In addition, elements can be defined to control the
attributes of other elements. This can be used, for example, to
simulate the action of control room ``knobs'' which vary the strengths
of multiple magnets in the lattice. Support ``I--beams'' can be
defined that determine the position of multiple element. 

To go beyond the things that \bmad can do, \bmad has been developed
to maximize flexibility. For example, different tracking methods may
be defined for a given element type and what the actual tracking
method used can be chosen at run time.


\vfill
\break
